---
title: "SycLLM User Study — Confirmatory Analysis"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r packages}
library(tidyverse)
library(psych)
library(knitr)
library(kableExtra)
library(effectsize)
library(emmeans)
library(rstatix)
```

# Data Loading & Cleaning

```{r load-data}
FILE_PATH <- "data/full_data.csv"
raw <- read_tsv(FILE_PATH, locale = locale(encoding = "UTF-16LE"), show_col_types = FALSE)
glimpse(raw)
```

```{r clean-data}
df <- raw %>%
  filter(FINISHED == 1, MODE != "admin", prolific_id != "test", STATUS == "complete") %>%
  mutate(
    age             = as.numeric(age),
    gender          = factor(gender, levels = 1:2, labels = c("male", "female")),
    model_condition = factor(model_condition,
                             levels = 1:3,
                             labels = c("unprompted", "sycophantic", "contrarian")),
    topic = factor(topic)
  )

cat("N after exclusions:", nrow(df), "\n")
```

```{r compute-scores}
row_mean <- function(df, cols) rowMeans(df[, cols], na.rm = TRUE)

df <- df %>%
  mutate(
    opinion_pre = case_when(
      topic == "space"   ~ space_02,
      topic == "quota"   ~ pre_quota,
      topic == "taxes"   ~ pre_taxes,
      topic == "animals" ~ pre_animals,
      topic == "speech"  ~ pre_speech
    ),
    opinion_post = case_when(
      topic == "space"   ~ post_space_04,
      topic == "quota"   ~ post_quota,
      topic == "taxes"   ~ post_taxes,
      topic == "animals" ~ post_animals,
      topic == "speech"  ~ post_speech
    ),
    opinion_change    = opinion_post - opinion_pre,
    confidence_change = post_confidence - pre_confidence,
    competence   = row_mean(., c("competence_01", "competence_02", "competence_03")),
    trust        = row_mean(., c("trust_01", "trust_02", "trust_03")),
    satisfaction = row_mean(., paste0("satisfaction_0", 1:5)),
    sycophancy   = row_mean(., c("sycophancy_01", "sycophancy_02")),
    credibility  = row_mean(., c("credibility_01", "credibility_02", "credibility_03")),
    continuance  = row_mean(., c("continuance_01", "continuance_02")),
    interaction_time = TIME_SUM / 60  # minutes
  )
```

# Sample Description

```{r sample-desc}
df %>%
  summarise(
    N = n(),
    Age_M = round(mean(age, na.rm = TRUE), 1),
    Age_SD = round(sd(age, na.rm = TRUE), 1),
    Age_min = min(age, na.rm = TRUE),
    Age_max = max(age, na.rm = TRUE)
  ) %>%
  kable(caption = "Sample demographics") %>%
  kable_styling(bootstrap_options = "striped")
```

```{r sample-gender}
df %>%
  count(gender) %>%
  mutate(pct = round(100 * n / sum(n), 1)) %>%
  kable(caption = "Gender distribution") %>%
  kable_styling(bootstrap_options = "striped")
```

```{r sample-conditions}
df %>%
  count(model_condition, topic) %>%
  pivot_wider(names_from = topic, values_from = n, values_fill = 0) %>%
  kable(caption = "Participants per condition x topic") %>%
  kable_styling(bootstrap_options = "striped")
```

# Scale Reliability

```{r reliability}
alpha_table <- tibble(
  Scale = c("Competence", "Trust", "Satisfaction", "Sycophancy",
            "Credibility", "Continuance"),
  Items = c(3, 3, 5, 2, 3, 2),
  Alpha = c(
    alpha(df[, c("competence_01", "competence_02", "competence_03")])$total$raw_alpha,
    alpha(df[, c("trust_01", "trust_02", "trust_03")])$total$raw_alpha,
    alpha(df[, paste0("satisfaction_0", 1:5)])$total$raw_alpha,
    alpha(df[, c("sycophancy_01", "sycophancy_02")])$total$raw_alpha,
    alpha(df[, c("credibility_01", "credibility_02", "credibility_03")])$total$raw_alpha,
    alpha(df[, c("continuance_01", "continuance_02")])$total$raw_alpha
  )
) %>%
  mutate(Alpha = round(Alpha, 3))

alpha_table %>%
  kable(caption = "Cronbach's alpha for composite scales") %>%
  kable_styling(bootstrap_options = "striped")
```

# Descriptive Statistics by Condition

```{r descriptives}
desc_vars <- c("opinion_pre", "opinion_post", "opinion_change",
               "pre_confidence", "post_confidence", "confidence_change",
               "competence", "trust", "satisfaction",
               "sycophancy", "credibility", "continuance", "interaction_time")

df %>%
  group_by(model_condition) %>%
  summarise(
    n = n(),
    across(all_of(desc_vars),
           list(M = ~round(mean(.x, na.rm = TRUE), 2),
                SD = ~round(sd(.x, na.rm = TRUE), 2)),
           .names = "{.col}_{.fn}")
  ) %>%
  pivot_longer(-model_condition, names_to = "Statistic", values_to = "Value") %>%
  pivot_wider(names_from = model_condition, values_from = Value) %>%
  kable(caption = "Descriptive statistics by condition") %>%
  kable_styling(bootstrap_options = "striped", font_size = 11) %>%
  scroll_box(height = "500px")
```

---

# Hypothesis Tests

## H1: Belief Reinforcement (Opinion Change)

> The sycophantic condition produces the strongest reinforcement of pre-existing beliefs, followed by the neutral condition, with the contrarian condition producing the weakest reinforcement.

```{r h1-paired-t}
for (cond in levels(df$model_condition)) {
  sub <- df %>% filter(model_condition == cond, !is.na(opinion_pre), !is.na(opinion_post))
  t_res <- t.test(sub$opinion_post, sub$opinion_pre, paired = TRUE)
  cat(sprintf(
    "%s (n=%d): M_pre=%.2f, M_post=%.2f, diff=%.2f, t(%d)=%.2f, p=%.3f\n",
    cond, nrow(sub),
    mean(sub$opinion_pre), mean(sub$opinion_post), mean(sub$opinion_change),
    t_res$parameter, t_res$statistic, t_res$p.value
  ))
}
```

```{r h1-plot, fig.width=7, fig.height=4}
df %>%
  filter(!is.na(opinion_change)) %>%
  ggplot(aes(x = model_condition, y = opinion_change, fill = model_condition)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.1, size = 2, alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "H1: Opinion change (post - pre) by condition",
       x = "Condition", y = "Opinion change") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

```{r h1-anova}
df_h1 <- df %>% filter(!is.na(opinion_change))

aov_h1 <- aov(opinion_change ~ model_condition, data = df_h1)
summary(aov_h1)

cat("\n--- Eta-squared ---\n")
eta_squared(aov_h1)

cat("\n--- Tukey HSD post-hoc ---\n")
TukeyHSD(aov_h1)

cat("\n--- Kruskal-Wallis (non-parametric) ---\n")
kw_h1 <- kruskal.test(opinion_change ~ model_condition, data = df_h1)
print(kw_h1)

cat("\n--- Dunn post-hoc ---\n")
df_h1 %>% dunn_test(opinion_change ~ model_condition, p.adjust.method = "bonferroni")
```

## H2: Confidence Change

> The sycophantic condition produces the greatest increase in confidence, followed by the neutral condition, with the contrarian condition producing the smallest increase (or a decrease).

```{r h2-paired-t}
for (cond in levels(df$model_condition)) {
  sub <- df %>% filter(model_condition == cond, !is.na(confidence_change))
  t_res <- t.test(sub$post_confidence, sub$pre_confidence, paired = TRUE)
  cat(sprintf(
    "%s (n=%d): M_pre=%.2f, M_post=%.2f, diff=%.2f, t(%d)=%.2f, p=%.3f\n",
    cond, nrow(sub),
    mean(sub$pre_confidence), mean(sub$post_confidence), mean(sub$confidence_change),
    t_res$parameter, t_res$statistic, t_res$p.value
  ))
}
```

```{r h2-plot, fig.width=7, fig.height=4}
df %>%
  filter(!is.na(confidence_change)) %>%
  ggplot(aes(x = model_condition, y = confidence_change, fill = model_condition)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.1, size = 2, alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "H2: Confidence change (post - pre) by condition",
       x = "Condition", y = "Confidence change") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

```{r h2-anova}
df_h2 <- df %>% filter(!is.na(confidence_change))

aov_h2 <- aov(confidence_change ~ model_condition, data = df_h2)
summary(aov_h2)

cat("\n--- Eta-squared ---\n")
eta_squared(aov_h2)

cat("\n--- Tukey HSD post-hoc ---\n")
TukeyHSD(aov_h2)

cat("\n--- Kruskal-Wallis (non-parametric) ---\n")
kw_h2 <- kruskal.test(confidence_change ~ model_condition, data = df_h2)
print(kw_h2)

cat("\n--- Dunn post-hoc ---\n")
df_h2 %>% dunn_test(confidence_change ~ model_condition, p.adjust.method = "bonferroni")
```

## H3: Trust

> Perceived trust differs across the three conditions.

```{r h3-test}
df_h3 <- df %>% filter(!is.na(trust))

aov_h3 <- aov(trust ~ model_condition, data = df_h3)
summary(aov_h3)
eta_squared(aov_h3)
TukeyHSD(aov_h3)

kw_h3 <- kruskal.test(trust ~ model_condition, data = df_h3)
print(kw_h3)
df_h3 %>% dunn_test(trust ~ model_condition, p.adjust.method = "bonferroni")
```

```{r h3-plot, fig.width=6, fig.height=4}
df_h3 %>%
  ggplot(aes(x = model_condition, y = trust, fill = model_condition)) +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.12, size = 2, alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "H3: Trust by condition", x = "Condition", y = "Trust (1-5)") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

## H4: Satisfaction

> User satisfaction differs across the three conditions.

```{r h4-test}
df_h4 <- df %>% filter(!is.na(satisfaction))

aov_h4 <- aov(satisfaction ~ model_condition, data = df_h4)
summary(aov_h4)
eta_squared(aov_h4)
TukeyHSD(aov_h4)

kw_h4 <- kruskal.test(satisfaction ~ model_condition, data = df_h4)
print(kw_h4)
df_h4 %>% dunn_test(satisfaction ~ model_condition, p.adjust.method = "bonferroni")
```

```{r h4-plot, fig.width=6, fig.height=4}
df_h4 %>%
  ggplot(aes(x = model_condition, y = satisfaction, fill = model_condition)) +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.12, size = 2, alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "H4: Satisfaction by condition", x = "Condition", y = "Satisfaction (1-5)") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

## H5: Credibility

> Perceived credibility differs across the three conditions.

```{r h5-test}
df_h5 <- df %>% filter(!is.na(credibility))

aov_h5 <- aov(credibility ~ model_condition, data = df_h5)
summary(aov_h5)
eta_squared(aov_h5)
TukeyHSD(aov_h5)

kw_h5 <- kruskal.test(credibility ~ model_condition, data = df_h5)
print(kw_h5)
df_h5 %>% dunn_test(credibility ~ model_condition, p.adjust.method = "bonferroni")
```

```{r h5-plot, fig.width=6, fig.height=4}
df_h5 %>%
  ggplot(aes(x = model_condition, y = credibility, fill = model_condition)) +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.12, size = 2, alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "H5: Credibility by condition", x = "Condition", y = "Credibility (1-5)") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

## H6: Competence

> Perceived competence differs across the three conditions.

```{r h6-test}
df_h6 <- df %>% filter(!is.na(competence))

aov_h6 <- aov(competence ~ model_condition, data = df_h6)
summary(aov_h6)
eta_squared(aov_h6)
TukeyHSD(aov_h6)

kw_h6 <- kruskal.test(competence ~ model_condition, data = df_h6)
print(kw_h6)
df_h6 %>% dunn_test(competence ~ model_condition, p.adjust.method = "bonferroni")
```

```{r h6-plot, fig.width=6, fig.height=4}
df_h6 %>%
  ggplot(aes(x = model_condition, y = competence, fill = model_condition)) +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.12, size = 2, alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "H6: Competence by condition", x = "Condition", y = "Competence (1-5)") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

## H7: Continuance

> Continuance interaction differs across the three conditions.

```{r h7-test}
df_h7 <- df %>% filter(!is.na(continuance))

aov_h7 <- aov(continuance ~ model_condition, data = df_h7)
summary(aov_h7)
eta_squared(aov_h7)
TukeyHSD(aov_h7)

kw_h7 <- kruskal.test(continuance ~ model_condition, data = df_h7)
print(kw_h7)
df_h7 %>% dunn_test(continuance ~ model_condition, p.adjust.method = "bonferroni")
```

```{r h7-plot, fig.width=6, fig.height=4}
df_h7 %>%
  ggplot(aes(x = model_condition, y = continuance, fill = model_condition)) +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.12, size = 2, alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "H7: Continuance by condition", x = "Condition", y = "Continuance (1-5)") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

## H8: Interaction Time

> Interaction time differs across the three conditions.

```{r h8-test}
df_h8 <- df %>% filter(!is.na(interaction_time))

aov_h8 <- aov(interaction_time ~ model_condition, data = df_h8)
summary(aov_h8)
eta_squared(aov_h8)
TukeyHSD(aov_h8)

kw_h8 <- kruskal.test(interaction_time ~ model_condition, data = df_h8)
print(kw_h8)
df_h8 %>% dunn_test(interaction_time ~ model_condition, p.adjust.method = "bonferroni")
```

```{r h8-plot, fig.width=6, fig.height=4}
df_h8 %>%
  ggplot(aes(x = model_condition, y = interaction_time, fill = model_condition)) +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.12, size = 2, alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "H8: Interaction time by condition", x = "Condition", y = "Time (minutes)") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

## H9: Manipulation Check — Perceived Sycophancy

> Users in the sycophantic condition report higher perceived sycophancy than those in the neutral and contrarian conditions.

```{r h9-desc}
df %>%
  group_by(model_condition) %>%
  summarise(
    n  = n(),
    M  = round(mean(sycophancy, na.rm = TRUE), 2),
    SD = round(sd(sycophancy, na.rm = TRUE), 2)
  ) %>%
  kable(caption = "Perceived sycophancy by condition") %>%
  kable_styling(bootstrap_options = "striped")
```

```{r h9-test}
df_h9 <- df %>% filter(!is.na(sycophancy))

aov_h9 <- aov(sycophancy ~ model_condition, data = df_h9)
summary(aov_h9)
eta_squared(aov_h9)
TukeyHSD(aov_h9)

kw_h9 <- kruskal.test(sycophancy ~ model_condition, data = df_h9)
print(kw_h9)
df_h9 %>% dunn_test(sycophancy ~ model_condition, p.adjust.method = "bonferroni")
```

```{r h9-plot, fig.width=6, fig.height=4}
df_h9 %>%
  ggplot(aes(x = model_condition, y = sycophancy, fill = model_condition)) +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.12, size = 2, alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "H9: Perceived sycophancy by condition",
       x = "Condition", y = "Sycophancy (1-5)") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

---

# Summary of All Hypothesis Tests

```{r summary-table}
# hyp_names <- c(
#   "H1: Opinion change", "H2: Confidence change", "H3: Trust",
#   "H4: Satisfaction", "H5: Credibility", "H6: Competence",
#   "H7: Continuance", "H8: Interaction time", "H9: Perceived sycophancy"
# )
# hyp_dvs <- c(
#   "opinion_change", "confidence_change", "trust",
#   "satisfaction", "credibility", "competence",
#   "continuance", "interaction_time", "sycophancy"
# )
# 
# summary_rows <- lapply(seq_along(hyp_dvs), function(i) {
#   dv <- hyp_dvs[i]
#   d <- df %>% filter(!is.na(.data[[dv]]))
#   aov_res <- aov(as.formula(paste(dv, "~ model_condition")), data = d)
#   aov_sum <- summary(aov_res)[[1]]
#   eta <- eta_squared(aov_res)
#   kw <- kruskal.test(as.formula(paste(dv, "~ model_condition")), data = d)
#   tibble(
#     Hypothesis = hyp_names[i],
#     F_val  = round(aov_sum$`F value`[1], 2),
#     p_anova = round(aov_sum$`Pr(>F)`[1], 3),
#     eta_sq = round(eta$Eta2[1], 3),
#     chi_sq = round(kw$statistic, 2),
#     p_kw   = round(kw$p.value, 3)
#   )
# })
# 
# bind_rows(summary_rows) %>%
#   kable(caption = "Summary of omnibus tests across all hypotheses",
#         col.names = c("Hypothesis", "F", "p (ANOVA)", "\u03b7\u00b2", "\u03c7\u00b2", "p (KW)")) %>%
#   kable_styling(bootstrap_options = "striped")
```

# Correlations Among Outcome Measures

```{r correlations, fig.width=7, fig.height=6}
df_cor <- df %>%
  select(opinion_change, confidence_change, competence, trust, satisfaction,
         sycophancy, credibility, continuance) %>%
  drop_na()

cor_mat <- cor(df_cor, use = "pairwise.complete.obs", method = "spearman")
round(cor_mat, 2) %>%
  kable(caption = "Spearman correlations among outcome measures") %>%
  kable_styling(bootstrap_options = "striped")
```
