---
title: "SycLLM — Chat Transcript Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 5)
```

```{r packages}
library(tidyverse)
library(jsonlite)
library(knitr)
library(kableExtra)
```

# Data Loading

```{r load-data}
raw_json <- fromJSON("all_chats.json", flatten = FALSE)

# Build a chat-level summary (one row per conversation)
chats <- tibble(
  case_id   = raw_json$case_id,
  case      = raw_json$case,
  condition = raw_json$condition,
  topic     = raw_json$topic,
  pre_opinion  = raw_json$pre_opinion,
  post_opinion = raw_json$post_opinion,
  transcript   = raw_json$transcript
)

# Derive per-chat metrics from transcripts
chats <- chats %>%
  mutate(
    n_messages    = map_int(transcript, nrow),
    # Exclude the system message for turn counting
    n_user_msgs   = map_int(transcript, ~ sum(.x$role == "user")),
    n_asst_msgs   = map_int(transcript, ~ sum(.x$role == "assistant")),
    n_turns       = n_user_msgs,  # each user message = 1 turn
    # Word counts
    user_words    = map_int(transcript, ~ sum(str_count(.x$content[.x$role == "user"], "\\S+"))),
    asst_words    = map_int(transcript, ~ sum(str_count(.x$content[.x$role == "assistant"], "\\S+"))),
    total_words   = user_words + asst_words,
    avg_user_msg_len = user_words / n_user_msgs,
    avg_asst_msg_len = asst_words / n_asst_msgs,
    # Opinion shift
    opinion_shift = post_opinion - pre_opinion
  )

chats <- chats %>%
  mutate(
    condition = factor(condition, levels = c("sycophantic", "unprompted", "contrarian")),
    topic     = factor(topic)
  )


mean(chats$n_turns)

# glimpse(chats %>% select(-transcript))
```


```{r summary-stats}
chats %>%
  group_by(condition) %>%
  summarise(
    n            = n(),
    mean_turns   = mean(n_turns),
    sd_turns     = sd(n_turns),
    mean_user_words = mean(user_words),
    mean_asst_words = mean(asst_words),
    mean_total_words = mean(total_words),
    .groups = "drop"
  ) %>%
  kable(digits = 1, caption = "Conversation metrics by condition") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Turn Count Analysis

```{r turns-by-condition, fig.cap="Number of user turns by condition"}
ggplot(chats, aes(x = condition, y = n_turns, fill = condition)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.5) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = "Condition", y = "Number of user turns",
       title = "User turns per conversation by condition") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r turns-anova}
turns_aov <- aov(n_turns ~ condition, data = chats)
summary(turns_aov)
```
-> number of conversation turns significantly differs across conditions
-> pairwise comparisons to find which conditions differ from each other

## Pairwise comparisions

```{r}
TukeyHSD(turns_aov)
```
The contrarian condition produces significantly fewer turns than the sycophantic condition (p = 0.003). The difference is about 2 fewer turns on average.


# Word Count Analysis

```{r wordcount-bars, fig.cap="Average word count by condition and role"}
chats %>%
  group_by(condition) %>%
  summarise(User = mean(user_words), Assistant = mean(asst_words), .groups = "drop") %>%
  pivot_longer(-condition, names_to = "role", values_to = "words") %>%
  ggplot(aes(x = condition, y = words, fill = role)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_fill_brewer(palette = "Paired") +
  labs(x = "Condition", y = "Mean total words per chat",
       title = "Total word count by condition and role") +
  theme_minimal()
```

```{r avg-msg-length, fig.cap="Average message length by condition"}
chats %>%
  select(condition, avg_user_msg_len, avg_asst_msg_len) %>%
  pivot_longer(-condition, names_to = "role", values_to = "words",
               names_pattern = "avg_(\\w+)_msg_len") %>%
  ggplot(aes(x = condition, y = words, fill = role)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_brewer(palette = "Paired", labels = c("Assistant", "User")) +
  labs(x = "Condition", y = "Average words per message",
       title = "Message length by condition and role") +
  theme_minimal()
```





```{r}


```










