---
title: "SycLLM — Chat Transcript Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 5)


# Plots Path
output_dir <- "/Users/maxeder/Documents/GitHub/ma_thesis_manuscript/Figures"
# 
# file.path(output_dir, "")
```

```{r packages}
library(tidyverse)
library(jsonlite)
library(knitr)
library(kableExtra)
library(dunn.test)
library(rstatix)
```

# Data Loading

```{r load-data}
raw_json <- fromJSON("all_chats.json", flatten = FALSE)

# Build a chat-level summary (one row per conversation)
chats <- tibble(
  case_id   = raw_json$case_id,
  case      = raw_json$case,
  condition = raw_json$condition,
  topic     = raw_json$topic,
  pre_opinion  = raw_json$pre_opinion,
  post_opinion = raw_json$post_opinion,
  transcript   = raw_json$transcript
)

# Derive per-chat metrics from transcripts
chats <- chats %>%
  mutate(
    n_messages    = map_int(transcript, nrow),
    # Exclude the system message for turn counting
    n_user_msgs   = map_int(transcript, ~ sum(.x$role == "user")),
    n_asst_msgs   = map_int(transcript, ~ sum(.x$role == "assistant")),
    n_turns       = n_user_msgs,  # each user message = 1 turn
    # Word counts
    user_words    = map_int(transcript, ~ sum(str_count(.x$content[.x$role == "user"], "\\S+"))),
    asst_words    = map_int(transcript, ~ sum(str_count(.x$content[.x$role == "assistant"], "\\S+"))),
    total_words   = user_words + asst_words,
    avg_user_msg_len = user_words / n_user_msgs,
    avg_asst_msg_len = asst_words / n_asst_msgs,
    # Opinion shift
    opinion_shift = post_opinion - pre_opinion
  )

chats <- chats %>%
  mutate(
    condition = factor(condition, levels = c("sycophantic", "unprompted", "contrarian")),
    topic     = factor(topic)
  )


mean(chats$n_turns)

# glimpse(chats %>% select(-transcript))
```


```{r summary-stats}
chats %>%
  group_by(condition) %>%
  summarise(
    n            = n(),
    mean_turns   = mean(n_turns),
    sd_turns     = sd(n_turns),
    mean_user_words = mean(user_words),
    mean_asst_words = mean(asst_words),
    mean_total_words = mean(total_words),
    .groups = "drop"
  ) %>%
  kable(digits = 1, caption = "Conversation metrics by condition") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Turn Count Analysis

```{r turns-by-condition, fig.cap="Number of user turns by condition"}
chats <- chats %>%
  mutate(
    condition = factor(condition, levels = c("contrarian","unprompted", "sycophantic"))
  )

ggplot(chats, aes(x = condition, y = n_turns, fill = condition)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.5) +
    scale_color_manual(values = colors) +
 scale_fill_manual(values = c("contrarian" = "#D4807A", "unprompted" = "#7BA7C9", "sycophantic" = "#82B88A")) +
  labs(x = "Condition", y = "Number of user turns",
       title = "User turns per conversation by condition") +
  theme_minimal() +
    scale_x_discrete(labels = c("contrarian" = "Contrarian", "unprompted" = "Unprompted", "sycophantic" = "Sycophantic")) +
  theme(legend.position = "none")
```


```{r}

chats <- chats %>%
  mutate(
    condition = factor(condition, levels = c("contrarian","unprompted", "sycophantic"))
  )


# Test
kw <-kruskal.test(n_turns ~ condition, data = chats)

kw

n <- nrow(chats)
epsilon_sq <- kw$statistic / (n - 1)
epsilon_sq
# Pairwise follow-up
dunn.test(chats$n_turns, chats$condition, method = "bh")

# Visualisation

# mean with cis
turns_plot <- ggplot(chats, aes(x = condition, y = n_turns,  color = condition)) +
  stat_summary(fun = mean, geom = "point", size = 3.5) +
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2, linewidth = 0.7) +
  labs(x = "Condition", y = "Number of Turns") +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.5, shape = 16) +
  scale_color_manual(values = c("contrarian" = "#D4807A", "unprompted" = "#7BA7C9", "sycophantic" = "#82B88A")) +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 10)) +
  scale_y_continuous(breaks = seq(0, 10, by = 2)) +
  scale_x_discrete(labels = c("contrarian" = "Contrarian", "unprompted" = "Unprompted", "sycophantic" = "Sycophantic")) +
  theme(legend.position = "none")

turns_plot
ggsave(file.path(output_dir, "turns_plot.pdf"), plot = turns_plot, width=8, height=5)
ggsave(file.path(output_dir, "turns_plot.png"), plot = turns_plot, width=8, height=5)
```




# Interaction Time

```{r}
# Load user study sosci dataset



FILE_PATH = "data/full_data.csv"
raw <- read_tsv(FILE_PATH, locale = locale(encoding = "UTF-16LE"),  show_col_types = FALSE)
df <- raw %>%
  filter(FINISHED == 1) %>%
  filter(TIME_SUM >= 300) %>%
  # drop admin / test entries
  filter(MODE != "admin", prolific_id != "test", FINISHED == 1, STATUS == "complete") %>%
  mutate(
    age             = as.numeric(age),
    gender          = factor(gender, levels = 1:2, labels = c("male", "female")),
    model_condition = factor(model_condition,
                             levels = 1:3,
                             labels = c("Unprompted", "Sycophantic", "Contrarian")),
    topic  = factor(topic)
  )



df <- df %>%
  mutate(
    model_condition = factor(model_condition, levels = c("Contrarian","Unprompted", "Sycophantic"))
  )

df$model_condition
df$TIME005 <- as.numeric(df$TIME005)

# Test
kw <- kruskal.test(TIME005 ~ model_condition, data = df)
kw
n <- nrow(df)
n
epsilon_sq <- kw$statistic / (n - 1)
epsilon_sq

# Pairwise follow-up
dunn.test(df$TIME005, df$model_condition, method = "holm")

# Visualisation

# mean with cis
time_plot <- ggplot(df, aes(x = model_condition, y = TIME005,  color = model_condition)) +
  stat_summary(fun = mean, geom = "point", size = 3.5) +
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2, linewidth = 0.7) +
  labs(x = "Condition", y = "Interaction Time in Seconds") +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.5, shape = 16) +
  scale_color_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
  theme_minimal() +
  # coord_cartesian(ylim = c(0, 10)) +
  theme(legend.position = "none")


time_plot
ggsave(file.path(output_dir, "time_plot.pdf"), plot = time_plot, width=8, height=5)
ggsave(file.path(output_dir, "time_plot.png"), plot = time_plot, width=8, height=5)

```




```{r turns-anova}
turns_aov <- aov(n_turns ~ condition, data = chats)
summary(turns_aov)
```
-> number of conversation turns significantly differs across conditions
-> pairwise comparisons to find which conditions differ from each other

## Pairwise comparisions

```{r}
TukeyHSD(turns_aov)
```
The contrarian condition produces significantly fewer turns than the sycophantic condition (p = 0.003). The difference is about 2 fewer turns on average.


<!-- # Word Count Analysis -->

<!-- ```{r wordcount-bars, fig.cap="Average word count by condition and role"} -->
<!-- chats %>% -->
<!--   group_by(condition) %>% -->
<!--   summarise(User = mean(user_words), Assistant = mean(asst_words), .groups = "drop") %>% -->
<!--   pivot_longer(-condition, names_to = "role", values_to = "words") %>% -->
<!--   ggplot(aes(x = condition, y = words, fill = role)) + -->
<!--   geom_col(position = "dodge", alpha = 0.8) + -->
<!--   scale_fill_brewer(palette = "Paired") + -->
<!--   labs(x = "Condition", y = "Mean total words per chat", -->
<!--        title = "Total word count by condition and role") + -->
<!--   theme_minimal() -->
<!-- ``` -->

```{r avg-msg-length, fig.cap="Average message length by condition"}
chats %>%
  select(condition, avg_user_msg_len, avg_asst_msg_len) %>%
  pivot_longer(-condition, names_to = "role", values_to = "words",
               names_pattern = "avg_(\\w+)_msg_len") %>%
  ggplot(aes(x = condition, y = words, fill = role)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_brewer(palette = "Paired", labels = c("Assistant", "User")) +
  labs(x = "Condition", y = "Average words per message",
       title = "Message length by condition and role") +
  theme_minimal()
```





```{r}


```










