
---
title: "SycLLM User Study — Analysis Confidence Change"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)


# Plots Path
output_dir <- "/Users/maxeder/Documents/GitHub/ma_thesis_manuscript/Figures"

# 
# file.path(output_dir, "")
```

# Packages

```{r packages}
library(tidyverse)
library(psych)
library(ggplot2)
library(knitr)
library(kableExtra)
library(lme4)
library(emmeans)
library(psych)
library(effectsize)
library(dunn.test)
library(dplyr)
library(jtools) 
library(ggalluvial)
library(ordinal)

```


# Data Loading & Cleaning


```{r load-data}

FILE_PATH = "data/full_data.csv"
raw <- read_tsv(FILE_PATH, locale = locale(encoding = "UTF-16LE"),  show_col_types = FALSE)

# Inspect
# glimpse(raw)
```

```{r clean-data}
df <- raw %>%
  filter(FINISHED == 1) %>%
  filter(TIME_SUM >= 300) %>%
  # drop admin / test entries
  filter(MODE != "admin", prolific_id != "test", FINISHED == 1, STATUS == "complete") %>%
  mutate(
    age             = as.numeric(age),
    gender          = factor(gender, levels = 1:2, labels = c("male", "female")),
    model_condition = factor(model_condition,
                             levels = 1:3,
                             labels = c("Unprompted", "Sycophantic", "Contrarian")),
    topic  = factor(topic)
  )

cat("N after exclusions:", nrow(df), "\n")
```

```{r attitude-scoring}
# Helper: row-wise mean ignoring NA
row_mean <- function(df, cols) {
  rowMeans(df[, cols], na.rm = TRUE)
}


topics <- c("space", "quota", "taxes", "animals", "speech")

# build composite pre-post opinion and confidence measures & experience composites via means
df <- df %>%
  mutate(
    # Pre Opinion
    opinion_pre = case_when(
      topic == "space"   ~ space_02,
      topic == "quota"   ~ pre_quota,
      topic == "taxes"   ~ pre_taxes,
      topic == "animals" ~ pre_animals,
      topic == "speech"  ~ pre_speech
    ),
    # Post-opinion
    opinion_post = case_when(
      topic == "space"   ~ post_space_04,
      topic == "quota"   ~ post_quota,
      topic == "taxes"   ~ post_taxes,
      topic == "animals" ~ post_animals,
      topic == "speech"  ~ post_speech
    ),
    opinion_change = opinion_post - opinion_pre,
    confidence_change = post_confidence - pre_confidence,
    # Scale composites
    competence  = row_mean(., c("competence_01", "competence_02", "competence_03")),
    trust       = row_mean(., c("trust_01",       "trust_02",       "trust_03")),
    satisfaction = row_mean(., paste0("satisfaction_0", 1:5)),
    sycophancy  = row_mean(., c("sycophancy_01",  "sycophancy_02")),
    credibility = row_mean(., c("credibility_01", "credibility_02", "credibility_03")),
    continuance = row_mean(., c("continuance_01", "continuance_02"))
  )
```



```{r}
df$model_condition <- factor(df$model_condition,
                             levels = c("Contrarian", "Unprompted", "Sycophantic"))
df$post_confidence <- factor(df$post_confidence, levels = 1:7, ordered = TRUE)

# Planned linear contrast = direct test of H2
contrasts(df$model_condition) <- contr.poly(3)

# Fit and inspect
m <- clm(post_confidence ~ model_condition + pre_confidence, data = df)
summary(m)
```

# FLow

```{r}
flow_colors <- c("Reinforced" = "#4A9E6B", "Unchanged" = "#C0C0C0", "Moderated" = "#C05050")
flow_alpha  <- c("Reinforced" = 0.75,      "Unchanged" = 0.75, "Moderated" = 0.75)

flow_graph_confidence <- df %>%
  filter(!is.na(pre_confidence), !is.na(post_confidence)) %>%
  mutate(
    confidence_pre_factor    = factor(pre_confidence,  levels = 7:1),
    confidence_post_factor   = factor(post_confidence, levels = 7:1),
    # model_condition       = factor(model_condition, levels = c("Unprompted", "Sycophantic", "Contrarian")),
    model_condition       = factor(model_condition, levels = c("Contrarian", "Unprompted", "Sycophantic")),
    change_type = case_when(
      post_confidence == pre_confidence                          ~ "Unchanged",
      post_confidence > pre_confidence               ~ "Reinforced", # more
      TRUE                                                                   ~ "Moderated"   # less
    ),
    change_type = factor(change_type, levels = c("Reinforced", "Unchanged", "Moderated"))
  ) %>%
  count(model_condition, confidence_pre_factor, confidence_post_factor, change_type) %>%
  group_by(model_condition) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup() %>%
  ggplot(aes(axis1 = confidence_pre_factor, axis2 = confidence_post_factor, y = prop)) +
  geom_alluvium(aes(fill = change_type, alpha = change_type), width = 1/6) +
  geom_stratum(fill = "white", colour = "grey30", width = 1/6) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3, fontface = "bold") +
  scale_x_discrete(limits = c("Pre-\nconfidence", "Post-\nconfidence"), expand = c(0.05, 0.05)) +
  scale_fill_manual(values = flow_colors, name = NULL) +
  scale_alpha_manual(values = flow_alpha, guide = "none") +
  guides(fill = guide_legend(override.aes = list(alpha = unname(flow_alpha)))) +
  facet_wrap(~model_condition) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percentage of participants") +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    # strip.background = element_rect(fill = NA, colour = "grey50", linewidth = 0),
    panel.grid.major.x = element_blank(), # Remove vertical lines
    panel.grid.minor = element_blank(),
    strip.text       = element_text(size = 11, face = "bold"),
    axis.text.x      = element_text(size = 10),
    axis.title.y = element_text(margin = margin(r = 10)),
    panel.spacing = unit(2, "lines")
  )



flow_graph_confidence
ggsave(file.path(output_dir, "flow_graph_confidence.pdf"), plot = flow_graph_confidence, width=9, height=5)
ggsave(file.path(output_dir, "flow_graph_confidence.png"), plot = flow_graph_confidence, width=9, height=5)
```





```{r}
em <- emmeans(m, ~ model_condition, mode = "mean.class")
em_df <- as.data.frame(em)

confidence_pred_mean <- ggplot(em_df, aes(x = model_condition, y = mean.class)) +
  geom_line(aes(group = 1), linetype = "dashed") +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL, color = model_condition), width = 0.1) +
  geom_point(aes(color = model_condition), size = 2.5) +
  labs(x = NULL, y = "Predicted post-confidence") +
  ylim(c(4, 7)) +
  scale_color_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none"
  )

confidence_pred_mean
ggsave(file.path(output_dir, "confidence_pred_mean.pdf"), plot = confidence_pred_mean, width=8, height=5)
ggsave(file.path(output_dir, "confidence_pred_mean.png"), plot = confidence_pred_mean, width=8, height=5)
```
Predicted mean post-confidence rating (1–7 scale) by model condition, estimated from a cumulative link mode. Points show expected ratings derived from the model's predicted probabilities; error bars are 95% confidence intervals. The dashed line visualizes the order tested by the linear contrast (Contrarian < Unprompted < Sycophantic). Y-axis truncated to emphasize between-condition differences.



```{r}

# ggplot(df, aes(x = model_condition, y = confidence_change)) +
#   geom_line(aes(group = 1), linetype = "dashed") +
#   geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL, color = model_condition), width = 0.1) +
#   geom_point(aes(color = model_condition), size = 2.5) +
#   labs(x = NULL, y = "Predicted post-confidence") +
#   ylim(c(4, 7)) +
#   scale_color_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
#   theme_minimal(base_size = 12) +
#   theme(
#     legend.position = "none"
#   )


ggplot(df, aes(x = model_condition, y = confidence_change, fill = model_condition)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  
  # suppress boxplot outliers when using jitter plot
  geom_boxplot(width = 0.4, fatten = 2, outlier.shape = NA) +
  # geom_violin() +
  geom_jitter(width = 0.15, size = 1.5, colour = "black", alpha = 0.4) +
  # suppress boxplot outliers when using jitter plot
  # geom_boxplot(width = 0.4, outlier.shape = NA) +
  
  # add mean
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  
  scale_fill_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
  labs(title = "Opinion change (post – pre) by condition",
       x = "Condition", y = "Attitude change") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```





# Moderators


```{r}
m_robust <- clm(post_confidence ~ model_condition + pre_confidence + 
                  ai_literacy + ai_usage, 
                data = df)
summary(m_robust)
```



```{r}

m_aov <- aov(confidence_change ~ model_condition, data = df)
summary.lm(m_aov) 
```

```{r}
summary(lm(confidence_change ~ model_condition * pre_confidence, data = df))
# Order the condition factor so the linear contrast points the right way:
# contrarian (-1) -> unprompted (0) -> sycophantic (+1)
df$model_condition <- factor(
  df$model_condition,
  levels = c("Contrarian", "Unprompted", "Sycophantic")
)

# Outcome must be an ORDERED factor for clm()
df$post_confidence <- factor(df$post_confidence, levels = 1:7, ordered = TRUE)
# Quick sanity checks
table(df$model_condition)
summary(df$pre_confidence)
table(df$model_condition, df$post_confidence)
 


# Linear contrast directly tests H2 (monotonic increase across the ordering).
# Quadratic contrast is a sanity check for non-monotonic patterns.
contrasts(df$model_condition) <- cbind(
  linear    = c(-1,  0,  1),
  quadratic = c( 1, -2,  1)
)
# Name the contrasts so they appear cleanly in output:
colnames(contrasts(df$model_condition)) <- c("linear", "quadratic")
 
# ---- 4. Fit the cumulative link model (ordinal ANCOVA) ----
# Controlling for pre_confidence handles baseline differences and
# regression to the mean. Treating pre as numeric is standard; if you
# prefer, you can enter it as an ordered factor with scored contrasts.
m <- clm(
  post_confidence ~ model_condition + pre_confidence,
  data = df,
  link = "logit"
)
summary(m)





# ---- 5. Primary test of H2: the linear contrast ----
# The row labeled "conditionlinear" in summary(m) IS the H2 test.
# Extract it cleanly:
coefs <- coef(summary(m))
h2_row <- coefs["model_conditionlinear", ]
h2_row
# Odds ratio + 95% CI for the linear contrast
est  <- h2_row["Estimate"]
se   <- h2_row["Std. Error"]
or   <- exp(est)
or_ci <- exp(est + c(-1, 1) * 1.96 * se)
cat(sprintf(
  "Linear trend: b = %.3f, SE = %.3f, OR = %.2f, 95%% CI [%.2f, %.2f], p = %.4f\n",
  est, se, or, or_ci[1], or_ci[2], h2_row["Pr(>|z|)"]
))
 

# ---- 7. Follow-up pairwise comparisons (exploratory) ----
emm <- emmeans(m, ~ model_condition, mode = "latent")
emm
# Pairwise on the latent scale with Holm correction
pairs(emm, adjust = "holm")
 
# ---- 8. Descriptives for the write-up ----
summary_ <- df %>%
  mutate(post_num = as.numeric(as.character(post_confidence))) %>%
  group_by(model_condition) %>%
  summarise(
    n         = n(),
    pre_mean  = mean(pre_confidence),
    pre_sd    = sd(pre_confidence),
    post_mean = mean(post_num),
    post_sd   = sd(post_num),
    post_med  = median(post_num),
    change_m  = mean(post_num - pre_confidence),
    change_sd = sd(post_num - pre_confidence)
  )
print(summary_)



df %>%
  mutate(
    pre  = pre_confidence,
    post = as.numeric(as.character(post_confidence))
  ) %>%
  tidyr::pivot_longer(c(pre, post), names_to = "time", values_to = "conf") %>%
  mutate(time = factor(time, levels = c("pre", "post"))) %>%
  ggplot(aes(x = time, y = conf, group = CASE)) +
  geom_line(alpha = 0.2) +
  stat_summary(aes(group = model_condition, color = model_condition),
               fun = mean, geom = "line", linewidth = 1.2) +
  stat_summary(aes(group = model_condition, color = model_condition),
               fun = mean, geom = "point", size = 3) +
  facet_wrap(~ model_condition) +
  labs(y = "Confidence (1-7)", x = NULL) +
  theme_minimal()



```

