
---
title: "SycLLM User Study — Analysis Reinforcement"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)


# Plots Path
output_dir <- "/Users/maxeder/Documents/GitHub/ma_thesis_manuscript/Figures"

# 
# file.path(output_dir, "")
```

# Packages

```{r packages}
library(tidyverse)
library(psych)
library(ggplot2)
library(knitr)
library(kableExtra)
library(lme4)
library(emmeans)
library(psych)
library(effectsize)
library(dunn.test)
library(dplyr)
library(jtools) 
library(ggalluvial)


```


# Data Loading & Cleaning


```{r load-data}

FILE_PATH = "data/full_data.csv"
raw <- read_tsv(FILE_PATH, locale = locale(encoding = "UTF-16LE"),  show_col_types = FALSE)

# Inspect
# glimpse(raw)
```

```{r clean-data}
df <- raw %>%
  filter(FINISHED == 1) %>%
  filter(TIME_SUM >= 300) %>%
  # drop admin / test entries
  filter(MODE != "admin", prolific_id != "test", FINISHED == 1, STATUS == "complete") %>%
  mutate(
    age             = as.numeric(age),
    gender          = factor(gender, levels = 1:2, labels = c("male", "female")),
    model_condition = factor(model_condition,
                             levels = 1:3,
                             labels = c("Unprompted", "Sycophantic", "Contrarian")),
    topic  = factor(topic)
  )

cat("N after exclusions:", nrow(df), "\n")
```

```{r attitude-scoring}
# Helper: row-wise mean ignoring NA
row_mean <- function(df, cols) {
  rowMeans(df[, cols], na.rm = TRUE)
}


topics <- c("space", "quota", "taxes", "animals", "speech")

# build composite pre-post opinion and confidence measures & experience composites via means
df <- df %>%
  mutate(
    # Pre Opinion
    opinion_pre = case_when(
      topic == "space"   ~ space_02,
      topic == "quota"   ~ pre_quota,
      topic == "taxes"   ~ pre_taxes,
      topic == "animals" ~ pre_animals,
      topic == "speech"  ~ pre_speech
    ),
    # Post-opinion
    opinion_post = case_when(
      topic == "space"   ~ post_space_04,
      topic == "quota"   ~ post_quota,
      topic == "taxes"   ~ post_taxes,
      topic == "animals" ~ post_animals,
      topic == "speech"  ~ post_speech
    ),
    opinion_change = opinion_post - opinion_pre,
    confidence_change = post_confidence - pre_confidence,
    # Scale composites
    competence  = row_mean(., c("competence_01", "competence_02", "competence_03")),
    trust       = row_mean(., c("trust_01",       "trust_02",       "trust_03")),
    satisfaction = row_mean(., paste0("satisfaction_0", 1:5)),
    sycophancy  = row_mean(., c("sycophancy_01",  "sycophancy_02")),
    credibility = row_mean(., c("credibility_01", "credibility_02", "credibility_03")),
    continuance = row_mean(., c("continuance_01", "continuance_02"))
  )
```


```{r}


df %>%
  filter(!is.na(pre_confidence), !is.na(post_confidence)) %>%
  mutate(
    confidence_pre_centered = round(pre_confidence) - 4,
    confidence_post_centered = round(post_confidence) - 4,
    changed = confidence_post_centered != confidence_pre_centered,
    abs_change = abs(confidence_post_centered - confidence_pre_centered)
  ) %>%
  # group_by(model_condition) %>%
  summarise(
    n = n(),
    n_changed = sum(changed),
    pct_changed = round(mean(changed) * 100, 1),
    mean_abs_change = round(mean(abs_change), 2),
    sd_abs_change = round(sd(abs_change), 2)
  )



df %>%
  filter(!is.na(pre_confidence), !is.na(post_confidence)) %>%
  mutate(
    confidence_pre_centered = round(pre_confidence) - 4,
    confidence_post_centered = round(post_confidence) - 4,
    changed = confidence_post_centered != confidence_pre_centered,
    abs_change = abs(confidence_post_centered - confidence_pre_centered)
  ) %>%
  group_by(model_condition) %>%
  summarise(
    n = n(),
    n_changed = sum(changed),
    pct_changed = round(mean(changed) * 100, 1),
    mean_abs_change = round(mean(abs_change), 2),
    sd_abs_change = round(sd(abs_change), 2)
  )

```


```{r flow-graph, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
# confidence reinforcement - Overview
# install.packages("ggalluvial")
# library(ggalluvial)

# Colors and alpha per change category
flow_colors <- c("Reinforced" = "#4A9E6B", "Unchanged" = "#C0C0C0", "Neither" = "#C0C0C0", "Moderated" = "#C05050")
flow_alpha  <- c("Reinforced" = 0.75,      "Unchanged" = 0.75,      "Neither" = 0.15,      "Moderated" = 0.75)

df %>%
  filter(!is.na(pre_confidence), !is.na(post_confidence)) %>%
  mutate(
    confidence_pre_centered  = round(pre_confidence)  - 4,   # recentre: -3 (disagree) to +3 (agree)
    confidence_post_centered = round(post_confidence) - 4,
    confidence_pre_factor    = factor(confidence_pre_centered,  levels = 3:-3),  # 3 on top, -3 on bottom
    confidence_post_factor   = factor(confidence_post_centered, levels = 3:-3),
    # model_condition       = factor(model_condition, levels = c("Unprompted", "Sycophantic", "Contrarian")),
    model_condition       = factor(model_condition, levels = c("Contrarian", "Unprompted", "Sycophantic")),
    change_type = case_when(
      confidence_pre_centered == 0                                              ~ "Neither",    # started neutral
      confidence_post_centered == confidence_pre_centered                          ~ "Unchanged",
      sign(confidence_post_centered) == sign(confidence_pre_centered) &
        abs(confidence_post_centered) > abs(confidence_pre_centered)               ~ "Reinforced", # same direction, more extreme
      TRUE                                                                   ~ "Moderated"   # toward centre, to 0, or crossed sides
    ),
    change_type = factor(change_type, levels = c("Reinforced", "Unchanged", "Neither", "Moderated"))
  ) %>%
  count(model_condition, confidence_pre_factor, confidence_post_factor, change_type) %>%
  group_by(model_condition) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup() %>%
  ggplot(aes(axis1 = confidence_pre_factor, axis2 = confidence_post_factor, y = prop)) +
  geom_alluvium(aes(fill = change_type, alpha = change_type), width = 1/6) +
  geom_stratum(fill = "white", colour = "grey30", width = 1/6) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3, fontface = "bold") +
  scale_x_discrete(limits = c("Pre-\nconfidence", "Post-\nconfidence"), expand = c(0.05, 0.05)) +
  scale_fill_manual(values = flow_colors, name = NULL) +
  scale_alpha_manual(values = flow_alpha, guide = "none") +
  guides(fill = guide_legend(override.aes = list(alpha = unname(flow_alpha)))) +
  facet_wrap(~model_condition) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percentage of participants") +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = NA, colour = "grey50", linewidth = 0),
    panel.grid.major.x = element_blank(), # Remove vertical lines
    panel.grid.minor = element_blank(),
    strip.text       = element_text(size = 11, face = "bold"),
    axis.text.x      = element_text(size = 10),
    axis.title.y = element_text(margin = margin(r = 10)),
    panel.spacing = unit(2, "lines")
  )
```



```{r reinforcement-means, fig.width=7, fig.height=4, message=FALSE, warning=FALSE}
# Confidence reinforcement - Graph
df %>%
  filter(!is.na(pre_confidence), !is.na(post_confidence)) %>%
  mutate(
    confidence_pre_centered  = round(pre_confidence)  - 4,
    confidence_post_centered = round(post_confidence) - 4,
    confidence_direction     = case_when(
      confidence_pre_centered < 0 ~ "Negative",
      confidence_pre_centered > 0 ~ "Positive",
      .default = "Neutral"
    ),
    confidence_reinforcement = (confidence_post_centered - confidence_pre_centered) * sign(confidence_pre_centered),
    # model_condition       = factor(model_condition, levels = c("Unprompted", "Sycophantic", "Contrarian"))
    model_condition       = factor(model_condition, levels = c("Contrarian", "Unprompted", "Sycophantic"))
  ) %>%
  filter(confidence_pre_centered != 0) %>% # Optional
  ggplot(aes(x = model_condition, y = confidence_reinforcement, color = model_condition)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_jitter(width = 0.15, height = 0.15, alpha = 0.4, size = 1.5, colour = "grey40", shape = 16) +
  scale_color_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
  # stat_summary(fun.data = mean_cl_normal, geom = "pointrange", size = 0.6, linewidth = 0.8) +
    stat_summary(
    fun.data = mean_cl_normal, 
    geom = "errorbar",    
    width = 0.2,          
    linewidth = 0.8
  ) +
  stat_summary(
    fun = mean, 
    geom = "point", 
    size = 3
  ) +
  labs(x = "Condition", y = "Confidence reinforcement") +
  theme_minimal(base_size = 12) + 
  theme(legend.position = "none")
```

```{r reinforcement-stats}
# Confidence reinforcement - Stats
df %>%
  filter(!is.na(pre_confidence), !is.na(post_confidence)) %>%
  mutate(
    confidence_pre_centered  = round(pre_confidence)  - 4,
    confidence_post_centered = round(post_confidence) - 4,
    confidence_reinforcement = (confidence_post_centered - confidence_pre_centered) * sign(confidence_pre_centered),
    # model_condition       = relevel(factor(model_condition), ref = "Unprompted")
  ) %>%
  filter(confidence_pre_centered != 0) %>%
lm(formula = confidence_reinforcement ~ model_condition + 0) %>% summary() # Testing all conditions against 0
```


# Moderators

```{r}
# Literacy
df$ai_literacy_c <- scale(df$ai_literacy, scale = FALSE)
df %>%
  filter(!is.na(pre_confidence), !is.na(post_confidence)) %>%
  mutate(
    confidence_pre_centered  = round(pre_confidence)  - 4,
    confidence_post_centered = round(post_confidence) - 4,
    confidence_reinforcement = (confidence_post_centered - confidence_pre_centered) * sign(confidence_pre_centered),
    # model_condition       = relevel(factor(model_condition), ref = "Unprompted")
  ) %>%
  filter(confidence_pre_centered != 0) %>%
lm(formula = confidence_reinforcement ~ model_condition * ai_literacy_c + 0) %>% summary()


# usage
df$ai_usage_c <- scale(df$ai_usage, scale = FALSE)

df %>%
  filter(!is.na(pre_confidence), !is.na(post_confidence)) %>%
  mutate(
    confidence_pre_centered  = round(pre_confidence)  - 4,
    confidence_post_centered = round(post_confidence) - 4,
    confidence_reinforcement = (confidence_post_centered - confidence_pre_centered) * sign(confidence_pre_centered),
    # model_condition       = relevel(factor(model_condition), ref = "Unprompted")
  ) %>%
  filter(confidence_pre_centered != 0) %>%
lm(formula = confidence_reinforcement ~ model_condition * ai_usage_c + 0) %>% summary()
```



# Confidence Change by Prior Confidence Across Model Conditions
 
```{r}


# center scales  
df$confidence_pre_centered <-  df$pre_confidence - 4
df$confidence_post_centered  <-  df$post_confidence - 4
df$confidence_change_cent  <-  df$confidence_post_centered - df$confidence_pre_centered


# +
#   scale_x_continuous(labels = c("-2", "-1\nCritical", "0\n|", "1\nAffirmative", "2")) + labs(x='Participant pre-task opinion', y='Opinion change toward model') 
# +
#   geom_label(x = -2, y = -2, hjust = 0, vjust = 0, size=3, aes(label = ifelse(model_stance=="Contra", "M = 0.215\nr = 0.103**", "M = 0.183\nr = -0.188***")))
df %>%
  mutate(
    model_condition = factor(model_condition, levels = c("Contrarian", "Unprompted", "Sycophantic"))
  ) %>%
ggplot(aes(y = confidence_change_cent, x = confidence_pre_centered, color = model_condition)) +
  geom_jitter(alpha = 0.5, size = 1, color = "black", width = 0.3, shape =16) +
  geom_smooth(method = "lm", alpha = 0.15) +
  stat_summary(fun.data = "mean_cl_normal", size = 0.4) +
  scale_color_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
  facet_grid(~ model_condition) +
  guides(color = "none") +
  labs(x='Pre-Interaction Confidence (Centered)', y='Confidence Change (Centered)') +
  # ylim(-3.3,3.3) +
  theme_minimal(base_size = 13) +
  theme(
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```
Changes in participant post-interaction confidence against model condition with a fitted linear regression line and 95% confidence bands.
Each point represents one participant. Colored points indicate the mean confidence change at each level of pre-interaction confidence with 95% confidence intervals. 


```{r}
# df %>%
# lm(formula = opinion_change_cent ~ model_condition * opinion_pre_cent) %>% summary()
# 
# # Contrarian as reference
# df %>%
#   mutate(model_condition = relevel(factor(model_condition), ref = "Contrarian")) %>%
#   lm(formula = opinion_change_cent ~ model_condition * opinion_pre_cent) %>% summary()
# 
# 
# # Sycophantic as reference
# df %>%
#   mutate(model_condition = relevel(factor(model_condition), ref = "Sycophantic")) %>%
#   lm(formula = opinion_change_cent ~ model_condition * opinion_pre_cent) %>% summary()
# 
# 
# # Unprompted as reference
# df %>%
#   mutate(model_condition = relevel(factor(model_condition), ref = "Unprompted")) %>%
#   lm(formula = opinion_change_cent ~ model_condition * opinion_pre_cent) %>% summary()


# All three conditions tested against zero
conditions <- c("Unprompted", "Sycophantic", "Contrarian")

lapply(conditions, function(ref) {
  df %>%
    mutate(model_condition = relevel(factor(model_condition), ref = ref)) %>%
    lm(formula = confidence_change_cent ~ model_condition * confidence_pre_centered) %>%
    summary() %>%
    coef() %>%
    as.data.frame() %>%
    rownames_to_column("term") %>%
    filter(term == "confidence_pre_centered") %>%
    mutate(reference = ref)
}) %>%
  bind_rows()
```



