---
title: "SycLLM User Study — Analysis Extremity"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)


# Plots Path
output_dir <- "/Users/maxeder/Documents/GitHub/ma_thesis_manuscript/Figures"

# 
# file.path(output_dir, "")
```

# Packages

```{r packages}
library(tidyverse)
library(psych)
library(ggplot2)
library(knitr)
library(kableExtra)
library(lme4)
library(emmeans)
library(psych)
library(effectsize)
library(dunn.test)
library(dplyr)
library(jtools) 
```

# Data Loading & Cleaning


```{r load-data}

FILE_PATH = "data/full_data.csv"
raw <- read_tsv(FILE_PATH, locale = locale(encoding = "UTF-16LE"),  show_col_types = FALSE)

# Inspect
# glimpse(raw)
```

```{r clean-data}
df <- raw %>%
  filter(FINISHED == 1) %>%
  filter(TIME_SUM >= 300) %>%
  # drop admin / test entries
  filter(MODE != "admin", prolific_id != "test", FINISHED == 1, STATUS == "complete") %>%
  mutate(
    age             = as.numeric(age),
    gender          = factor(gender, levels = 1:2, labels = c("male", "female")),
    model_condition = factor(model_condition,
                             levels = 1:3,
                             labels = c("Unprompted", "Sycophantic", "Contrarian")),
    topic  = factor(topic)
  )

cat("N after exclusions:", nrow(df), "\n")
```

```{r attitude-scoring}
# Helper: row-wise mean ignoring NA
row_mean <- function(df, cols) {
  rowMeans(df[, cols], na.rm = TRUE)
}


topics <- c("space", "quota", "taxes", "animals", "speech")

# build composite pre-post opinion and confidence measures & experience composites via means
df <- df %>%
  mutate(
    # Pre Opinion
    opinion_pre = case_when(
      topic == "space"   ~ space_02,
      topic == "quota"   ~ pre_quota,
      topic == "taxes"   ~ pre_taxes,
      topic == "animals" ~ pre_animals,
      topic == "speech"  ~ pre_speech
    ),
    # Post-opinion
    opinion_post = case_when(
      topic == "space"   ~ post_space_04,
      topic == "quota"   ~ post_quota,
      topic == "taxes"   ~ post_taxes,
      topic == "animals" ~ post_animals,
      topic == "speech"  ~ post_speech
    ),
    opinion_change = opinion_post - opinion_pre,
    confidence_change = post_confidence - pre_confidence,
    # Scale composites
    competence  = row_mean(., c("competence_01", "competence_02", "competence_03")),
    trust       = row_mean(., c("trust_01",       "trust_02",       "trust_03")),
    satisfaction = row_mean(., paste0("satisfaction_0", 1:5)),
    sycophancy  = row_mean(., c("sycophancy_01",  "sycophancy_02")),
    credibility = row_mean(., c("credibility_01", "credibility_02", "credibility_03")),
    continuance = row_mean(., c("continuance_01", "continuance_02"))
  )
```



# Opinion Extremity

```{r}

# # Data Prep
# 
# df$condition <- relevel(factor(df$model_condition), ref = "Unprompted")
# df$extremityPre <-  abs(4 - as.numeric(df$opinion_pre))
# df$extremityPost  <-  abs(4 - as.numeric(df$opinion_post))
# 
# 
# # OLS
# 
# fit_ols <- function(dat, formula) {
#   m <- lm(formula, data = dat)
#   coef_tab <- summary(m)$coefficients
#   ci_tab <- confint(m)
#   combined <- cbind(coef_tab, ci_tab)
#   em <- emmeans(m, ~ condition, data = dat)
#   list(model = m, coef_table = combined, emmeans = em)
# }
# 
# res <- fit_ols(df, extremityPost ~ condition + extremityPre)
# 

# # Contrasts
# 
# res
# 
# custom_contrasts <- contrast(res$emmeans, method = list(
#   "Agreeable vs Neutral"      = c(1, 0, -1),
#   "Agreeable vs Disagreeable" = c(1, -1, 0),
#   "Disagreeable vs Neutral"   = c(0, 1, -1)
# ), adjust = "none")
# summary(custom_contrasts, infer = c(TRUE, TRUE))
# 
# 
# # Coef Table 
# 
# to_coef_tbl <- function(res, keep_intercept = FALSE) {
#   df <- as.data.frame(res$coef_table) |>
#     tibble::rownames_to_column("term")
#   
#   if (!keep_intercept) df <- df |> filter(term != "(Intercept)")
#   
#   df |>
#     mutate(
#       Term = recode(term,
#         "conditionAgreeable"    = "Agreeable",
#         "conditionDisagreeable" = "Disagreeable",
#         "extremityPre"          = "Pre extremity"
#       ),
#       CI = sprintf("[%.3f, %.3f]", `2.5 %`, `97.5 %`)
#     ) |>
#     transmute(Term, Estimate = round(Estimate, 3),
#               SE = round(`Std. Error`, 3), t = round(`t value`, 3),
#               p = round(`Pr(>|t|)`, 3), CI)
# }
# 
# to_coef_tbl(res)
# 





# Data Prep

df$condition <- relevel(factor(df$model_condition), ref = "Unprompted")
df$extremityPre <-  abs(4 - as.numeric(df$opinion_pre))
df$extremityPost  <-  abs(4 - as.numeric(df$opinion_post))


# OLS

fit_ols <- function(dat, formula) {
  m <- lm(formula, data = dat)
  coef_tab <- summary(m)$coefficients
  ci_tab <- confint(m)
  combined <- cbind(coef_tab, ci_tab)
  em <- emmeans(m, ~ condition)
  list(model = m, coef_table = combined, emmeans = em)
}

res <- fit_ols(df, extremityPost ~ condition + extremityPre)

res


# Plot

coef_df <- as.data.frame(res$coef_table) |>
  tibble::rownames_to_column("term") |>
  filter(grepl("^condition", term)) |>
  mutate(
    term_label = recode(term,
      "conditionSycophantic"    = "Sycophantic",
      "conditionContrarian" = "Contrarian"
    ),
    term_label = factor(term_label, levels = c("Sycophantic", "Contrarian"))
  )

ext_plot <- ggplot(coef_df, aes(x = Estimate, y = term_label, color = term_label)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = `2.5 %`, xmax = `97.5 %`), height = 0.15) +
  labs(x = "Change in Opinion Extremity (0-3 Scale)", y = NULL) +
  scale_x_continuous(limits=c(-1, 1)) +
  scale_color_manual(values = c("Sycophantic" = "#82B88A",
                                 "Contrarian" = "#D4807A")) +
  theme_apa() +
  theme(
    panel.grid.major.x = element_line(color = "grey90"), 
    panel.grid.minor.x = element_line(color = "grey90"),
    legend.position = "none",
  )
ext_plot
# ggsave(file.path(output_dir, "opinion_extremity.pdf"), plot = ext_plot, width=8, height=5)
# ggsave(file.path(output_dir, "opinion_extremity.png"), plot = ext_plot, width=8, height=5)
```





# Opinion Change by Prior Opinion Across Model Conditions
 
```{r}


# center scales  
df$opinion_pre_cent <-  df$opinion_pre - 4
df$opinion_post_cent  <-  df$opinion_post - 4
df$opinion_change_cent  <-  df$opinion_post_cent - df$opinion_pre_cent


# +
#   scale_x_continuous(labels = c("-2", "-1\nCritical", "0\n|", "1\nAffirmative", "2")) + labs(x='Participant pre-task opinion', y='Opinion change toward model') 
# +
#   geom_label(x = -2, y = -2, hjust = 0, vjust = 0, size=3, aes(label = ifelse(model_stance=="Contra", "M = 0.215\nr = 0.103**", "M = 0.183\nr = -0.188***")))
df %>%
  mutate(
    model_condition = factor(model_condition, levels = c("Contrarian", "Unprompted", "Sycophantic"))
  ) %>%
ggplot(aes(y = opinion_change_cent, x = opinion_pre_cent, color = model_condition)) +
  geom_jitter(alpha = 0.5, size = 1, color = "black", width = 0.3, shape =16) +
  geom_smooth(method = "lm", alpha = 0.15) +
  stat_summary(fun.data = "mean_cl_normal", size = 0.4) +
  scale_color_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
  facet_grid(~ model_condition) +
  guides(color = "none") +
  labs(x='Pre-Interaction Opinion (Centered)', y='Opinion Change (Centered)') +
  ylim(-3.3,3.3) +
  theme_minimal(base_size = 13) +
  theme(
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```
Changes in participant post-interaction opinion against model condition with a fitted linear regression line and 95% confidence bands.
Each point represents one participant. Red points indicate the mean opinion change at each level of pre-interaction opinion with 95% confidence intervals. The dashed horizontal line marks zero opinion change.




```{r}
t.test(df$extremityPost[df$condition == "Sycophantic"],
       df$extremityPre[df$condition == "Sycophantic"],
       paired = TRUE)

t.test(df$extremityPost[df$condition == "Unprompted"],
       df$extremityPre[df$condition == "Unprompted"],
       paired = TRUE)

t.test(df$extremityPost[df$condition == "Contrarian"],
       df$extremityPre[df$condition == "Contrarian"],
       paired = TRUE)
```

WIP

As additional exploratory analysis, paired t-tests within each condition were conducted to test whether post-interaction extremity differed from pre-interaction extremity.  The sycophantic condition showed a non-significant trend toward increased extremity (mean change = 0.29, p = .062), while the unprompted (mean change = 0.10, p = .540) and contrarian (mean change = 0.04, p = .806) conditions showed no meaningful change.



However, without a no-conversation control condition, these within-group comparisons cannot distinguish genuine attitude change from artifacts such as regression to the mean or repeated measurement effects. These preliminary findings motivate the inclusion of a control condition in future studies.........







---------------------------







# Confidence Extremity 


```{r}



# Data Prep

df$condition <- relevel(factor(df$model_condition), ref = "Unprompted")
df$extremityConfPre <-  abs(4 - as.numeric(df$pre_confidence))
df$extremityConfPost  <-  abs(4 - as.numeric(df$post_confidence))


# OLS

fit_ols <- function(dat, formula) {
  m <- lm(formula, data = dat)
  coef_tab <- summary(m)$coefficients
  ci_tab <- confint(m)
  combined <- cbind(coef_tab, ci_tab)
  em <- emmeans(m, ~ condition)
  list(model = m, coef_table = combined, emmeans = em)
}

res_conf <- fit_ols(df, extremityConfPost ~ condition + extremityConfPre)

res_conf


# Plot

coef_conf_df <- as.data.frame(res_conf$coef_table) |>
  tibble::rownames_to_column("term") |>
  filter(grepl("^condition", term)) |>
  mutate(
    term_label = recode(term,
      "conditionSycophantic"    = "Sycophantic",
      "conditionContrarian" = "Contrarian"
    ),
    term_label = factor(term_label, levels = c("Sycophantic", "Contrarian"))
  )

ext_plot_conf <- ggplot(coef_conf_df, aes(x = Estimate, y = term_label, color = term_label)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = `2.5 %`, xmax = `97.5 %`), height = 0.15) +
  labs(x = "Change in Opinion Extremity (0-3 Scale)", y = NULL) +
  scale_x_continuous(limits=c(-1.5, 1.5)) +
  scale_color_manual(values = c("Sycophantic" = "#82B88A",
                                 "Contrarian" = "#D4807A")) +
  theme_apa() +
  theme(
    panel.grid.major.x = element_line(color = "grey90"), 
    panel.grid.minor.x = element_line(color = "grey90"),
    legend.position = "none",
  )
ext_plot_conf
# ggsave(file.path(output_dir, "confidence_extremity.pdf"), plot = ext_plot, width=8, height=5)
# ggsave(file.path(output_dir, "confidence_extremity.png"), plot = ext_plot, width=8, height=5)
```

```{r}
t.test(df$extremityConfPost[df$condition == "Sycophantic"],
       df$extremityConfPre[df$condition == "Sycophantic"],
       paired = TRUE)

t.test(df$extremityConfPost[df$condition == "Unprompted"],
       df$extremityConfPre[df$condition == "Unprompted"],
       paired = TRUE)

t.test(df$extremityConfPost[df$condition == "Contrarian"],
       df$extremityConfPre[df$condition == "Contrarian"],
       paired = TRUE)
```

WIP
text ref above



# Confidence Change by Prior Opinion Across Model Conditions
 
```{r}


# center scales  
df$conf_pre_cent <-  df$pre_confidence - 4
df$conf_post_cent  <-  df$post_confidence - 4
df$conf_change_cent  <-  df$conf_post_cent - df$conf_pre_cent


# +
#   scale_x_continuous(labels = c("-2", "-1\nCritical", "0\n|", "1\nAffirmative", "2")) + labs(x='Participant pre-task opinion', y='Opinion change toward model') 
# +
#   geom_label(x = -2, y = -2, hjust = 0, vjust = 0, size=3, aes(label = ifelse(model_stance=="Contra", "M = 0.215\nr = 0.103**", "M = 0.183\nr = -0.188***")))
df %>%
  mutate(
    model_condition = factor(model_condition, levels = c("Contrarian", "Unprompted", "Sycophantic"))
  ) %>%
ggplot(aes(y = conf_change_cent, x = conf_pre_cent, color = model_condition)) +
  geom_jitter(alpha = 0.5, size = 1, color = "black", width = 0.3, shape =16) +
  geom_smooth(method = "lm", alpha = 0.15) +
  stat_summary(fun.data = "mean_cl_normal", size = 0.4) +
  scale_color_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
  facet_grid(~ model_condition) +
  labs(x='Pre-Interaction Confidence (Centered)', y='Confidence Change (Centered)') +
  guides(color = "none") +
  ylim(-3.3,3.3) +
  theme_minimal(base_size = 13) +
  theme(
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```
