---
title: "SycLLM User Study — Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)


# Plots Path
output_dir <- "/Users/maxeder/Documents/GitHub/ma_thesis_manuscript/Figures"

# 
# file.path(output_dir, "")
```

# Packages

```{r packages}
library(tidyverse)
library(psych)
library(ggplot2)
library(knitr)
library(kableExtra)
library(lme4)
library(emmeans)
library(psych)
library(effectsize)
library(dunn.test)
```

# Data Loading & Cleaning


```{r load-data}

FILE_PATH = "data/full_data.csv"
raw <- read_tsv(FILE_PATH, locale = locale(encoding = "UTF-16LE"),  show_col_types = FALSE)

# Inspect
glimpse(raw)
```

```{r clean-data}
df <- raw %>%
  filter(FINISHED == 1) %>%
  filter(TIME_SUM >= 300) %>%
  # drop admin / test entries
  filter(MODE != "admin", prolific_id != "test", FINISHED == 1, STATUS == "complete") %>%
  mutate(
    age             = as.numeric(age),
    gender          = factor(gender, levels = 1:2, labels = c("male", "female")),
    model_condition = factor(model_condition,
                             levels = 1:3,
                             labels = c("Unprompted", "Sycophantic", "Contrarian")),
    topic  = factor(topic)
  )

cat("N after exclusions:", nrow(df), "\n")
```

```{r attitude-scoring}
# Helper: row-wise mean ignoring NA
row_mean <- function(df, cols) {
  rowMeans(df[, cols], na.rm = TRUE)
}


topics <- c("space", "quota", "taxes", "animals", "speech")

# build composite pre-post opinion and confidence measures & experience composites via means
df <- df %>%
  mutate(
    # Pre Opinion
    opinion_pre = case_when(
      topic == "space"   ~ space_02,
      topic == "quota"   ~ pre_quota,
      topic == "taxes"   ~ pre_taxes,
      topic == "animals" ~ pre_animals,
      topic == "speech"  ~ pre_speech
    ),
    # Post-opinion
    opinion_post = case_when(
      topic == "space"   ~ post_space_04,
      topic == "quota"   ~ post_quota,
      topic == "taxes"   ~ post_taxes,
      topic == "animals" ~ post_animals,
      topic == "speech"  ~ post_speech
    ),
    opinion_change = opinion_post - opinion_pre,
    confidence_change = post_confidence - pre_confidence,
    # Scale composites
    competence  = row_mean(., c("competence_01", "competence_02", "competence_03")),
    trust       = row_mean(., c("trust_01",       "trust_02",       "trust_03")),
    satisfaction = row_mean(., paste0("satisfaction_0", 1:5)),
    sycophancy  = row_mean(., c("sycophancy_01",  "sycophancy_02")),
    credibility = row_mean(., c("credibility_01", "credibility_02", "credibility_03")),
    continuance = row_mean(., c("continuance_01", "continuance_02"))
  )
```


------------------------------------------------------------------------------------------------------------------------------------


# Cronbach's Alpha for Experience Scales

```{r}

scale_items <- list(
   competence  = c("competence_01", "competence_02", "competence_03"),
    trust       = c("trust_01",       "trust_02",       "trust_03"),
    satisfaction = paste0("satisfaction_0", 1:5),
    sycophancy  = c("sycophancy_01",  "sycophancy_02"),
    credibility = c("credibility_01", "credibility_02", "credibility_03"),
    continuance = c("continuance_01", "continuance_02")
)

# Run alpha for each scale
# alphas <- lapply(scale_items, function(vars) {
#   items <- df %>% select(all_of(vars))
#   psych::alpha(items)
# })

# Inspect e.g.:
# alphas$competence


alpha_values <- sapply(scale_items, function(vars) {
  items <- df %>% select(all_of(vars))
  psych::alpha(items)$total$raw_alpha
})

alpha_values
```





------------------------------------------------------------------------------------------------------------------------------------


# Opinion Change

## Overall opinion change


```{r}
# Wilcoxon signed-rank test

wilcox.test(df$opinion_pre, df$opinion_post, paired = TRUE)

median(df$opinion_pre)
median(df$opinion_post)
mean(df$opinion_post - df$opinion_pre)

# Effect size (rank-biserial correlation)
rank_biserial(df$opinion_pre, df$opinion_post, paired = TRUE)
```
A Wilcoxon signed-rank test indicated that post-interaction opinion scores were significantly higher than pre-interaction scores (V = 571, p = .024, r = 0.33).


## Pre–Post within Each Condition

```{r att-change-tests}
for (cond in levels(df$model_condition)) {
  sub <- df %>% filter(model_condition == cond, !is.na(opinion_pre), !is.na(opinion_post))
  t_res <- t.test(sub$opinion_post, sub$opinion_pre, paired = TRUE)
  cat(sprintf(
    "\n%s (n=%d): M_pre=%.2f, M_post=%.2f, diff=%.2f, t(%d)=%.2f, p=%.3f\n",
    cond, nrow(sub),
    mean(sub$opinion_pre), mean(sub$opinion_post), mean(sub$opinion_change),
    t_res$parameter, t_res$statistic, t_res$p.value
  ))
}
  
```


## Opinion Change by Condition

```{r att-change-bar, fig.width=6, fig.height=4}
df %>%
  mutate(
    model_condition = factor(model_condition, levels = c("Contrarian", "Unprompted", "Sycophantic"))
  ) %>%
  ggplot(aes(x = model_condition, y = opinion_change, fill = model_condition)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  
  # suppress boxplot outliers when using jitter plot
  geom_boxplot(width = 0.4, fatten = 2, outlier.shape = NA) +
  # geom_violin() +
  geom_jitter(width = 0.15, size = 1.5, colour = "black", alpha = 0.4) +
  # suppress boxplot outliers when using jitter plot
  # geom_boxplot(width = 0.4, outlier.shape = NA) +
  
  # add mean
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  
  scale_fill_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
  labs(title = "Opinion change (post – pre) by condition",
       x = "Condition", y = "Attitude change") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```




# ```{r}
# 
# library(ggridges)
# df %>%
#   mutate(
#     model_condition = factor(model_condition, levels = c("Contrarian", "Unprompted", "Sycophantic"))
#   ) %>%
#   ggplot(aes(y = model_condition, x = opinion_change, fill = model_condition)) +
#   geom_density_ridges() +
#   scale_fill_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
#   theme_ridges() + 
#   theme(legend.position = "none")
# ```



## ANOVA / Kruskal-Wallis on Opinion Change

```{r}

# hist(df$opinion_change)

# ANOVA
# aov_op <- aov(opinion_change ~ model_condition, data = df)
# print(summary(aov_op))

# Kruskal
# kw_op <- kruskal.test(opinion_change ~ model_condition, data = df)
# kw_op
# 
# # maybe report kruskal due to non-normality?
# print(median(df$opinion_change))
# print(quantile(df$opinion_change, .25))
# print(quantile(df$opinion_change, .75))
# 



kw_op <- kruskal.test(opinion_change ~ model_condition, data = df)
kw_op <- kruskal.test(opinion_change ~ model_condition, data = df)

n <- nrow(df %>% filter(!is.na(opinion_change)))
epsilon_sq <- kw_op$statistic / (n - 1)

cat(sprintf("chi-squared(%.0f)=%.2f  p=%.3f  ε²=%.3f\n",
            kw_op$parameter, kw_op$statistic, kw_op$p.value, epsilon_sq))

desc_op <- df %>%
  filter(!is.na(opinion_change)) %>%
  group_by(model_condition) %>%
  summarise(mdn = median(opinion_change),
            q1  = quantile(opinion_change, .25),
            q3  = quantile(opinion_change, .75),
            .groups = "drop")

cat(sprintf("  %-12s Mdn=%.2f  IQR=[%.2f, %.2f]\n",
                desc_op$model_condition, desc_op$mdn, desc_op$q1, desc_op$q3))


summary(kw_op)
names(df)
df %>% mutate(opinion_pre = opinion_pre - 4, opinion_post = opinion_post - 4) %>%
lm(formula = opinion_change ~ model_condition +opinion_pre + model_condition : opinion_pre) %>%
  summary()




# df %>% 
#   # center opinion scale
#   mutate(opinion_pre = opinion_pre - 4, opinion_post = opinion_post - 4) %>%
#   mutate(extremity_pre = abs(opinion_pre), extremity_post = abs(opinion_post), 
#          extremity_change = extremity_post - extremity_pre) %>%
# lm(formula = extremity_change ~ model_condition) %>%
#   summary()
```

## Extremity Approach


```{r}
df %>% 
  # center opinion scale
  mutate(opinion_pre = opinion_pre - 4, opinion_post = opinion_post - 4) %>%
  mutate(extremity_pre = abs(opinion_pre), extremity_post = abs(opinion_post), 
         extremity_change = extremity_post - extremity_pre) %>%
lm(formula = extremity_change ~ model_condition) %>%
  summary()


# model <- lm(GunControlExtremityPost ~ condition + GunControlExtremityPre, data = data)
df %>% 
  # center opinion scale
  mutate(opinion_pre = opinion_pre - 4, opinion_post = opinion_post - 4) %>%
  mutate(extremity_pre = abs(opinion_pre), extremity_post = abs(opinion_post), 
         extremity_change = extremity_post - extremity_pre) %>%
lm(formula = extremity_change ~ model_condition) %>%
  summary()


df %>% 
  # center opinion scale
  mutate(opinion_pre = opinion_pre - 4, opinion_post = opinion_post - 4) %>%
  mutate(extremity_pre = abs(opinion_pre), extremity_post = abs(opinion_post), 
         extremity_change = extremity_post - extremity_pre) %>%
lm(formula = extremity_post ~ model_condition + extremity_pre) %>%
  summary()
```

```{r}
library(broom)

df %>%
  mutate(opinion_pre = opinion_pre - 4, opinion_post = opinion_post - 4,
         extremity_pre = abs(opinion_pre), extremity_post = abs(opinion_post),
         extremity_change = extremity_post - extremity_pre) %>%
  lm(formula = extremity_change ~ model_condition, data = .) %>%
  tidy(conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip() +
  labs(x = "", y = "Estimated Effect on Extremity Change") +
  theme_minimal()
```

```{r}
df %>%
   mutate(opinion_pre = opinion_pre - 4, opinion_post = opinion_post - 4,
         extremity_pre = abs(opinion_pre), extremity_post = abs(opinion_post),
         extremity_change = extremity_post - extremity_pre) %>%
  lm(formula = extremity_change ~ model_condition, data = .) %>%
  tidy(conf.int = TRUE) %>%
  mutate(term = str_remove(term, "model_condition"),
         term = ifelse(term == "(Intercept)", "unprompted", term)) %>%
  ggplot(aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip() +
  labs(x = "", y = "Estimated Effect on Extremity Change") +
  theme_minimal()
```

## Bavel / Rathje Reference Code
## Extremity (Pre-Registered)


```{r}

# Recode Variables
data$GunControlExtremityPost <- abs(50 - as.numeric(data$GunControlPost_1))
data$GunControlExtremityPre <- abs(50 - as.numeric(data$GunControlSupportPre_1))

# Fit Model
model <- lm(GunControlExtremityPost ~ condition + GunControlExtremityPre, data = data)

# Robust SEs and confidence intervals
robust_se_output <- coeftest(model, vcov = vcovHC(model, type = "HC2"))
combined_output <- cbind(robust_se_output, coefci(model, vcov. = vcovHC(model, type = "HC2")))
round(combined_output, 3)

# Robust covariance matrix for emmeans
robust_cov <- vcovHC(model, type = "HC2")

# Estimated marginal means with robust SEs
em <- emmeans(model, ~ condition, vcov. = robust_cov)

# Planned contrasts with robust SEs and CIs
custom_contrasts <- contrast(em, method = list(
  "Agreeable vs Neutral" = c(0, 1, 0, -1),
  "Agreeable vs Disagreeable" = c(0, 1, -1, 0),
  "Disagreeable vs Neutral" = c(0, 0, 1, -1)
), adjust = "none")

summary(custom_contrasts, infer = c(TRUE, TRUE))

blue_color <- wes_palette("Zissou1", 5)[1]
wes_palette("Zissou1")
emm_df <- as.data.frame(em)
emm_df$condition <- factor(
  emm_df$condition,
  levels = c("Control",
             "Agreeable", 
             "Disagreeable", 
             "Unprompted")
)

#Graph for Plotting
ExtremityPlot <- ggplot(emm_df, aes(x = condition, y = emmean, color = condition)) +
  geom_point(size = 4, color = blue_color) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.1, size = 1, color = blue_color) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    legend.position = "none"
  ) +
  labs(
    title = "Extremity",
    y = "Estimated Marginal Mean",
    x = "Condition"
  )

```
## adjusted code from bavel paper
```{r}

# Recode Variables
# data$GunControlExtremityPost <- abs(50 - as.numeric(data$GunControlPost_1))
# data$GunControlExtremityPre <- abs(50 - as.numeric(data$GunControlSupportPre_1))

# (opinion_pre = opinion_pre - 4, opinion_post = opinion_post - 4,

install.packages("lmtest")
install.packages("sandwich")  # for robust/HAC standard errors
install.packages("wesanderson")

library(lmtest)
library(sandwich)
library("wesanderson")
  
df$ext_pre <-  abs(4 - as.numeric(df$opinion_pre))
df$ext_post  <-  abs(4 - as.numeric(df$opinion_post))

# Fit Model
model <- lm(ext_post ~ model_condition + ext_pre, data = df)

# Robust SEs and confidence intervals
robust_se_output <- coeftest(model, vcov = vcovHC(model, type = "HC2"))
combined_output <- cbind(robust_se_output, coefci(model, vcov. = vcovHC(model, type = "HC2")))
round(combined_output, 3)

# Robust covariance matrix for emmeans
robust_cov <- vcovHC(model, type = "HC2")

# Estimated marginal means with robust SEs
em <- emmeans(model, ~ model_condition, vcov. = robust_cov)

# Planned contrasts with robust SEs and CIs
# custom_contrasts <- contrast(em, method = list(
#   "Agreeable vs Neutral" = c(0, 1, 0, -1),
#   "Agreeable vs Disagreeable" = c(0, 1, -1, 0),
#   "Disagreeable vs Neutral" = c(0, 0, 1, -1)
# ), adjust = "none")
# 
# summary(custom_contrasts, infer = c(TRUE, TRUE))
# 


blue_color <- wes_palette("Zissou1", 5)[1]
wes_palette("Zissou1")
emm_df <- as.data.frame(em)
print(emm_df$model_condition)
emm_df$model_condition <- factor(
  emm_df$model_condition,
  levels = c(
             "Unprompted", 
             "Sycophantic", 
             "Contrarian")
)

#Graph for Plotting
ExtremityPlot <- ggplot(emm_df, aes(x = model_condition, y = emmean, color = model_condition)) +
  geom_point(size = 4, color = blue_color) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.1, size = 1, color = blue_color) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    legend.position = "none"
  ) +
  labs(
    title = "Extremity",
    y = "Estimated Marginal Mean",
    x = "Condition"
  )
ExtremityPlot


ggplot(emm_df, aes(x = model_condition, y = emmean, ymin = lower.CL, ymax = upper.CL)) +
  geom_pointrange() +
  theme_minimal() +
  labs(
    title = "Certainty",
    y = "Estimated Mean", x = "Condition"
  )
```








## adjusted code plot

```{r}
data$condition <- relevel(factor(data$condition), ref = "Unprompted")


```











```{r}
pos <- position_jitter(width=.3, height=.0, seed=1)
df %>% 
  mutate(opinion_pre = opinion_pre - 4, opinion_post = opinion_post - 4) %>%
  ggplot(aes(x=opinion_pre, y=opinion_change)) + facet_grid(~model_condition) + ylim(-3.3,3.3) + geom_jitter(width=.3, height=0.1, color="grey") + geom_smooth(method = "lm", color="blue") +
  stat_summary(fun="mean", color="red") + stat_summary(fun.data="mean_cl_normal", color="red")
  

  geom_smooth(se=FALSE)
hist(df$opinion_pre, breaks=50)
table(df$opinion_pre, useNA = "always")
?geom_smooth
```




------------------------------------------------------------------------------------------------------------------------------------

# Confidence Change


## Overall confidence change


```{r}
# Wilcoxon signed-rank test

wilcox.test(df$pre_confidence, df$post_confidence, paired = TRUE)

median(df$pre_confidence)
median(df$post_confidence)
mean(df$post_confidence - df$pre_confidence)

# Effect size (rank-biserial correlation)
rank_biserial(df$post_confidence, df$pre_confidence, paired = TRUE)
```
A Wilcoxon signed-rank test indicated that post-interaction confidence scores were significantly higher than pre-interaction scores (V = 730.5, p = .022, r = 0.32).



## Pre–Post within Each Condition

```{r att-change-tests}
for (cond in levels(df$model_condition)) {
  sub <- df %>% filter(model_condition == cond, !is.na(pre_confidence), !is.na(post_confidence))
  t_res <- t.test(sub$post_confidence, sub$pre_confidence, paired = TRUE)
  cat(sprintf(
    "\n%s (n=%d): M_pre=%.2f, M_post=%.2f, diff=%.2f, t(%d)=%.2f, p=%.3f\n",
    cond, nrow(sub),
    mean(sub$pre_confidence), mean(sub$post_confidence), mean(sub$confidence_change),
    t_res$parameter, t_res$statistic, t_res$p.value
  ))
}

```



## Confidence Change by Condition

```{r att-change-bar, fig.width=6, fig.height=4}
df %>%
  mutate(
    model_condition = factor(model_condition, levels = c("Contrarian", "Unprompted", "Sycophantic"))
  ) %>%
  ggplot(aes(x = model_condition, y = confidence_change, fill = model_condition)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  
  # suppress boxplot outliers when using jitter plot
  geom_boxplot(width = 0.4, fatten = 2, outlier.shape = NA) +
  # geom_jitter(width = 0.1, size = 2, colour = "black", alpha = 0.4) +
  geom_jitter(width = 0.15, size = 1.5, colour = "black", alpha = 0.4) +
  # suppress boxplot outliers when using jitter plot
  # geom_boxplot(width = 0.4, outlier.shape = NA) +
  
  # add mean
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  
  scale_fill_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
  labs(title = "Confidence change (post – pre) by condition",
       x = "Condition", y = "Attitude change") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

## ANOVA / Kruskal-Wallis on Confidence Change

```{r}
# 
# # hist(df$confidence_change)
# 
# # ANOVA
# aov_conf <- aov(confidence_change ~ model_condition, data = df)
# print(summary(aov_conf))
# 
# # Kruskal-Wallis 
# kw_conf <- kruskal.test(confidence_change ~ model_condition, data = df)
# print(kw_conf)
# 
# 
# # maybe report kruskal due to non-normality?
# 
# # pairwise follow-ups if significant
# # TukeyHSD(aov_conf)


kw_conf <- kruskal.test(confidence_change ~ model_condition, data = df)

n <- nrow(df %>% filter(!is.na(confidence_change)))
epsilon_sq <- kw_conf$statistic / (n - 1)

cat(sprintf("chi-squared(%.0f)=%.2f  p=%.3f  ε²=%.3f\n",
            kw_conf$parameter, kw_conf$statistic, kw_conf$p.value, epsilon_sq))

desc_conf <- df %>%
  filter(!is.na(confidence_change)) %>%
  group_by(model_condition) %>%
  summarise(mdn = median(confidence_change),
            q1  = quantile(confidence_change, .25),
            q3  = quantile(confidence_change, .75),
            .groups = "drop")

cat(sprintf("  %-12s Mdn=%.2f  IQR=[%.2f, %.2f]\n",
                desc_conf$model_condition, desc_conf$mdn, desc_conf$q1, desc_conf$q3))

```






------------------------------------------------------------------------------------------------------------------------------------


# Experience Measures

```{r outcomes-long}
outcomes <- c("competence", "trust", "satisfaction", "sycophancy",
              "credibility", "continuance")

df_out <- df %>%
  select(CASE, model_condition, all_of(outcomes)) %>%
  pivot_longer(cols = all_of(outcomes),
               names_to = "outcome", values_to = "value") %>%
  filter(!is.na(value))
```

```{r outcomes-plot, fig.width=9, fig.height=6}
df_out %>%
    mutate(
    model_condition = factor(model_condition, levels = c("Contrarian", "Unprompted", "Sycophantic")),
    outcome = tools::toTitleCase(outcome)
  ) %>%
  ggplot(aes(x = model_condition, y = value, fill = model_condition)) +
  # geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_boxplot(width = 0.4, , outlier.color = "grey50", outlier.size = 0.7) +
  
  # add mean
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  # suppress boxplot outliers when using jitter plot
  # geom_boxplot(width = 0.4, outlier.shape = NA) +
  # geom_jitter(width = 0.12, size = 1.5, alpha = 0.3) +
  # scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  # scale_fill_manual(values = c("firebrick", "steelblue", "firebrick")) +
  scale_fill_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
  facet_wrap(~outcome, scales = "free_y") +
  labs(title = "Outcome measures by condition",
       x = "Condition", y = "Score", fill = "Condition") +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 25, hjust = 1),
    legend.position = "none",
    strip.background = element_rect(fill = NA, colour = "grey50", linewidth = 1),
    strip.text = element_text(size = 10)
  )
```

```{r outcomes-kw}

# 
# for (out in outcomes) {
#   sub <- df %>% filter(!is.na(.data[[out]]))
#   kw <- kruskal.test(as.formula(paste(out, "~ model_condition")), data = sub)
#   cat(sprintf("%-15s  chi-squared(%.0f)=%.2f  p=%.3f\n",
#               out, kw$parameter, kw$statistic, kw$p.value))
#   if (kw$p.value < 0.05) {
#     cat("  Dunn test pairwise comparisons:\n")
#     dunn.test(sub[[out]], sub$model_condition, method = "bh")
# 
#   }
#   cat("\n")
#   cat("\n")
#   cat("\n")
# }



# for (out in outcomes) {
#   sub <- df %>% filter(!is.na(.data[[out]]))
#   kw <- kruskal.test(as.formula(paste(out, "~ model_condition")), data = sub)
#   
#   cat(sprintf("%-15s  chi-squared(%.0f)=%.2f  p=%.3f\n",
#               out, kw$parameter, kw$statistic, kw$p.value))
#   
#   desc <- sub %>%
#     group_by(model_condition) %>%
#     summarise(mdn = median(.data[[out]]),
#               q1  = quantile(.data[[out]], .25),
#               q3  = quantile(.data[[out]], .75),
#               .groups = "drop")
#   
#   for (i in seq_len(nrow(desc))) {
#     cat(sprintf("  %-12s Mdn=%.2f  IQR=[%.2f, %.2f]\n",
#                 desc$model_condition[i], desc$mdn[i], desc$q1[i], desc$q3[i]))
#   }
#   
#   if (kw$p.value < 0.05) {
#     cat("  Dunn test pairwise comparisons:\n")
#     dunn.test(sub[[out]], sub$model_condition, method = "bh")
#   }
#   cat("\n\n\n")
# }


for (out in outcomes) {
  sub <- df %>% filter(!is.na(.data[[out]]))
  kw <- kruskal.test(as.formula(paste(out, "~ model_condition")), data = sub)
  
  n <- nrow(sub)
  epsilon_sq <- kw$statistic / (n - 1)
  
  cat(sprintf("%-15s  chi-squared(%.0f)=%.2f  p=%.3f  ε²=%.3f\n",
              out, kw$parameter, kw$statistic, kw$p.value, epsilon_sq))
  
  desc <- sub %>%
    group_by(model_condition) %>%
    summarise(mdn = median(.data[[out]]),
              q1  = quantile(.data[[out]], .25),
              q3  = quantile(.data[[out]], .75),
              .groups = "drop")
  
  for (i in seq_len(nrow(desc))) {
    cat(sprintf("  %-12s Mdn=%.2f  IQR=[%.2f, %.2f]\n",
                desc$model_condition[i], desc$mdn[i], desc$q1[i], desc$q3[i]))
  }
  
  if (kw$p.value < 0.05) {
    cat("  Dunn test pairwise comparisons:\n")
    dunn.test(sub[[out]], sub$model_condition, method = "bh")
  }
  cat("\n\n\n")
}

# print(dunn.test(df$trust, df$model_condition, method = "bh"))


```

<!-- # Perceived Sycophancy -->


<!-- ```{r sycophancy-check} -->
<!-- df %>% -->
<!--   group_by(model_condition) %>% -->
<!--   summarise( -->
<!--     n = n(), -->
<!--     M = round(mean(sycophancy, na.rm = TRUE), 2), -->
<!--     SD = round(sd(sycophancy, na.rm = TRUE), 2) -->
<!--   ) %>% -->
<!--   kable(caption = "Perceived sycophancy by condition") %>% -->
<!--   kable_styling(bootstrap_options = "striped") -->
<!-- ``` -->



# Correlations Among Outcome Measures

```{r correlations, fig.width=7, fig.height=6}
df_cor <- df %>%
  select(opinion_change, competence, trust, satisfaction,
         sycophancy, credibility, continuance) %>%
  drop_na()

cor_mat <- cor(df_cor, use = "pairwise.complete.obs", method = "spearman")
round(cor_mat, 2) %>%
  kable(caption = "Spearman correlations among outcome measures") %>%
  kable_styling(bootstrap_options = "striped")
```




-----------------------------------------------------------------------------------------------------------------------------------



# Qualitative Results

## Randomized Qualitative Experience Question

```{r qual-results}

labels <- c(
  "A043" = "Trust",
  "A059" = "Sycophancy",
  "A060" = "Credibility",
  "A045" = "Satisfaction",
  "A042" = "Competence",
  "A061" = "Continuance"
)

# df %>%
#   summarise(qualQ = labels[qualQ],
#             if (qualQ == "Competence") competence) %>%
#   kable(caption = "Open Answers") %>%
#   kable_styling(bootstrap_options = "striped")


df %>%
  mutate(
    qualitative_question = labels[qualQ],
    question_mean = case_match(
      qualQ,
      "A043" ~ round(trust, 1),
      "A059" ~ round(sycophancy, 1),
      "A060" ~ round(credibility, 1),
      "A045" ~ round(satisfaction, 1),
      "A042" ~ round(competence, 1),
      "A061" ~ round(continuance, 1),
      
      .default = NA
    )
  ) %>%
  select(CASE, qualitative_question, question_mean, expQual) %>%
  arrange(qualitative_question) %>%
  kable(caption = "Qualitative Experience Question") %>%
  kable_styling(bootstrap_options = "striped")
```



## Purpose of study

```{r}

df %>%
  select(CASE, purpose) %>%
  kable(caption = "Purpose Question") %>%
  kable_styling(bootstrap_options = "striped")
```

## Optional Debrief Question

```{r}
df %>%
  filter(!is.na(debrief_question)) %>%
  select(CASE, debrief_question) %>%
  kable(caption = "Debrief Question") %>%
  kable_styling(bootstrap_options = "striped")
```

