---
title: "SycLLM User Study — Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Packages

```{r packages}
library(tidyverse)
library(psych)
library(ggplot2)
library(knitr)
library(kableExtra)
library(lme4)
library(emmeans)
library(psych)
```

# Data Loading & Cleaning


```{r load-data}

FILE_PATH = "data/full_data.csv"
raw <- read_tsv(FILE_PATH, locale = locale(encoding = "UTF-16LE"),  show_col_types = FALSE)

# Inspect
glimpse(raw)
```

```{r clean-data}
df <- raw %>%
  filter(FINISHED == 1) %>%
  filter(TIME_SUM >= 300) %>%
  # drop admin / test entries
  filter(MODE != "admin", prolific_id != "test", FINISHED == 1, STATUS == "complete") %>%
  mutate(
    age             = as.numeric(age),
    gender          = factor(gender, levels = 1:2, labels = c("male", "female")),
    model_condition = factor(model_condition,
                             levels = 1:3,
                             labels = c("unprompted", "sycophantic", "contrarian")),
    topic  = factor(topic)
  )

cat("N after exclusions:", nrow(df), "\n")
```

```{r attitude-scoring}
# Helper: row-wise mean ignoring NA
row_mean <- function(df, cols) {
  rowMeans(df[, cols], na.rm = TRUE)
}


topics <- c("space", "quota", "taxes", "animals", "speech")

# build composite pre-post opinion and confidence measures & experience composites via means
df <- df %>%
  mutate(
    # Pre Opinion
    opinion_pre = case_when(
      topic == "space"   ~ space_02,
      topic == "quota"   ~ pre_quota,
      topic == "taxes"   ~ pre_taxes,
      topic == "animals" ~ pre_animals,
      topic == "speech"  ~ pre_speech
    ),
    # Post-opinion
    opinion_post = case_when(
      topic == "space"   ~ post_space_04,
      topic == "quota"   ~ post_quota,
      topic == "taxes"   ~ post_taxes,
      topic == "animals" ~ post_animals,
      topic == "speech"  ~ post_speech
    ),
    opinion_change = opinion_post - opinion_pre,
    confidence_change = post_confidence - pre_confidence,
    # Scale composites
    competence  = row_mean(., c("competence_01", "competence_02", "competence_03")),
    trust       = row_mean(., c("trust_01",       "trust_02",       "trust_03")),
    satisfaction = row_mean(., paste0("satisfaction_0", 1:5)),
    sycophancy  = row_mean(., c("sycophancy_01",  "sycophancy_02")),
    credibility = row_mean(., c("credibility_01", "credibility_02", "credibility_03")),
    continuance = row_mean(., c("continuance_01", "continuance_02"))
  )
```





# Cronbach's Alpha for Experience Scales

```{r}

scale_items <- list(
   competence  = c("competence_01", "competence_02", "competence_03"),
    trust       = c("trust_01",       "trust_02",       "trust_03"),
    satisfaction = paste0("satisfaction_0", 1:5),
    sycophancy  = c("sycophancy_01",  "sycophancy_02"),
    credibility = c("credibility_01", "credibility_02", "credibility_03"),
    continuance = c("continuance_01", "continuance_02")
)

# Run alpha for each scale
# alphas <- lapply(scale_items, function(vars) {
#   items <- df %>% select(all_of(vars))
#   psych::alpha(items)
# })

# Inspect e.g.:
# alphas$competence


alpha_values <- sapply(scale_items, function(vars) {
  items <- df %>% select(all_of(vars))
  psych::alpha(items)$total$raw_alpha
})

alpha_values
```

# Opinion Change

## Pre–Post within Each Condition

```{r att-change-tests}
for (cond in levels(df$model_condition)) {
  sub <- df %>% filter(model_condition == cond, !is.na(opinion_pre), !is.na(opinion_post))
  t_res <- t.test(sub$opinion_post, sub$opinion_pre, paired = TRUE)
  cat(sprintf(
    "\n%s (n=%d): M_pre=%.2f, M_post=%.2f, diff=%.2f, t(%d)=%.2f, p=%.3f\n",
    cond, nrow(sub),
    mean(sub$opinion_pre), mean(sub$opinion_post), mean(sub$opinion_change),
    t_res$parameter, t_res$statistic, t_res$p.value
  ))
}
  
#   
# for (t in levels(df$topic)) {
#   sub <- df %>% filter(topic == t, !is.na(opinion_pre), !is.na(opinion_post))
#   if (nrow(sub) < 2) next
#   t_res <- t.test(sub$opinion_post, sub$opinion_pre, paired = TRUE)
#   cat(sprintf(
#     "\n%s (n=%d): M_pre=%.2f, M_post=%.2f, diff=%.2f, t(%d)=%.2f, p=%.3f\n",
#     cond, nrow(sub),
#     mean(sub$opinion_pre), mean(sub$opinion_post), mean(sub$opinion_change),
#     t_res$parameter, t_res$statistic, t_res$p.value
#   ))
# }
```

## 6.2 Attitude Change by Condition

```{r att-change-plot, fig.width=7, fig.height=4}
df_att <- df %>%
  filter(!is.na(opinion_pre), !is.na(opinion_post)) %>%
  pivot_longer(cols = c(opinion_pre, opinion_post),
               names_to = "time", values_to = "attitude") %>%
  mutate(time = factor(time, levels = c("opinion_pre", "opinion_post"),
                       labels = c("Pre", "Post")))

df_att %>%
  ggplot(aes(x = time, y = attitude, colour = model_condition, group = CASE)) +
  geom_line(alpha = 0.4) +
  geom_point(size = 2) +
  stat_summary(aes(group = model_condition), fun = mean,
               geom = "line", linewidth = 1.2, linetype = "dashed") +
  facet_wrap(~model_condition) +
  scale_colour_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "Individual attitude trajectories by condition",
       x = "Time", y = "Attitude (1–5)", colour = "Condition") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```


## Opinion Change by Condition

```{r att-change-bar, fig.width=6, fig.height=4}
df %>%
  filter(!is.na(opinion_change)) %>%
  ggplot(aes(x = model_condition, y = opinion_change, fill = model_condition)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  # geom_violin() + 
  geom_jitter(width = 0.1, size = 2, colour = "black", alpha = 0.6) +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  scale_y_continuous(limits = c(-3, 3)) +
  labs(title = "Attitude change (post – pre) by condition",
       x = "Condition", y = "Attitude change") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

## ANOVA / Kruskal-Wallis on Opinion Change

```{r att-change-anova}
df_opinion_change <- df %>% filter(!is.na(opinion_change))

if (nrow(df_opinion_change) >= 5) {
  aov_res <- aov(opinion_change ~ model_condition, data = df_opinion_change)
  print(summary(aov_res))

  # Non-parametric alternative given small N
  kw <- kruskal.test(opinion_change ~ model_condition, data = df_opinion_change)
  cat(sprintf("\nKruskal-Wallis: chi²(%.0f) = %.2f, p = %.3f\n",
              kw$parameter, kw$statistic, kw$p.value))
}
```


# Confidence Change


## Pre–Post within Each Condition

```{r att-change-tests}
for (cond in levels(df$model_condition)) {
  sub <- df %>% filter(model_condition == cond, !is.na(pre_confidence), !is.na(post_confidence))
  t_res <- t.test(sub$post_confidence, sub$pre_confidence, paired = TRUE)
  cat(sprintf(
    "\n%s (n=%d): M_pre=%.2f, M_post=%.2f, diff=%.2f, t(%d)=%.2f, p=%.3f\n",
    cond, nrow(sub),
    mean(sub$pre_confidence), mean(sub$post_confidence), mean(sub$confidence_change),
    t_res$parameter, t_res$statistic, t_res$p.value
  ))
}

```



## Confidence Change by Condition

```{r att-change-bar, fig.width=6, fig.height=4}
df %>%
  filter(!is.na(confidence_change)) %>%
  ggplot(aes(x = model_condition, y = confidence_change, fill = model_condition)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_boxplot(alpha = 0.7, width = 0.4, fatten = 2) +
  geom_jitter(width = 0.1, size = 2, colour = "black", alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "white") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, colour = "black") +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "Confidence change (post – pre) by condition",
       x = "Condition", y = "Attitude change") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

## Anova

```{r}
# overall test
aov_model <- aov(confidence_change ~ model_condition, data = df)
summary(aov_model)

# pairwise follow-ups if significant
TukeyHSD(aov_model)
```



## Krusakl Walis on Confidence Change

<!-- TBD -->





# Experience Measures

```{r outcomes-long}
outcomes <- c("competence", "trust", "satisfaction", "sycophancy",
              "credibility", "continuance")

df_out <- df %>%
  select(CASE, model_condition, all_of(outcomes)) %>%
  pivot_longer(cols = all_of(outcomes),
               names_to = "outcome", values_to = "value") %>%
  filter(!is.na(value))
```

```{r outcomes-plot, fig.width=9, fig.height=6}
df_out %>%
  ggplot(aes(x = model_condition, y = value, fill = model_condition)) +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.12, size = 1.5, alpha = 0.3) +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  facet_wrap(~outcome, scales = "free_y") +
  labs(title = "Outcome measures by condition",
       x = "Condition", y = "Score", fill = "Condition") +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 25, hjust = 1),
        legend.position = "none")
```

```{r outcomes-kw}
for (out in outcomes) {
  sub <- df %>% filter(!is.na(.data[[out]]))
  if (length(unique(sub$model_condition)) < 2 || nrow(sub) < 4) next
  kw <- kruskal.test(as.formula(paste(out, "~ model_condition")), data = sub)
  cat(sprintf("%-15s  chi²(%.0f)=%.2f  p=%.3f\n",
              out, kw$parameter, kw$statistic, kw$p.value))
}
```

# 8. Perceived Sycophancy

Does perceived sycophancy (`sycophancy_01/02`) differ between conditions?

```{r sycophancy-check}
df %>%
  group_by(model_condition) %>%
  summarise(
    n = n(),
    M = round(mean(sycophancy, na.rm = TRUE), 2),
    SD = round(sd(sycophancy, na.rm = TRUE), 2)
  ) %>%
  kable(caption = "Perceived sycophancy by condition") %>%
  kable_styling(bootstrap_options = "striped")
```

```{r sycophancy-plot, fig.width=5, fig.height=4}
df %>%
  filter(!is.na(sycophancy)) %>%
  ggplot(aes(x = model_condition, y = sycophancy, fill = model_condition)) +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.12, size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "Perceived sycophancy by condition",
       x = "Condition", y = "Sycophancy (1–5)") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

# 9. Correlations Among Outcome Measures

```{r correlations, fig.width=7, fig.height=6}
df_cor <- df %>%
  select(opinion_change, competence, trust, satisfaction,
         sycophancy, credibility, continuance) %>%
  drop_na()

cor_mat <- cor(df_cor, use = "pairwise.complete.obs", method = "spearman")
round(cor_mat, 2) %>%
  kable(caption = "Spearman correlations among outcome measures") %>%
  kable_styling(bootstrap_options = "striped")
```

# 10. Qualitative Results

## Randomized Qualitative Experience Question

```{r qual-results}

labels <- c(
  "A043" = "Trust",
  "A059" = "Sycophancy",
  "A060" = "Credibility",
  "A045" = "Satisfaction",
  "A042" = "Competence",
  "A061" = "Continuance"
)

# df %>%
#   summarise(qualQ = labels[qualQ],
#             if (qualQ == "Competence") competence) %>%
#   kable(caption = "Open Answers") %>%
#   kable_styling(bootstrap_options = "striped")


df %>%
  mutate(
    qualitative_question = labels[qualQ],
    question_mean = case_match(
      qualQ,
      "A043" ~ round(trust, 1),
      "A059" ~ round(sycophancy, 1),
      "A060" ~ round(credibility, 1),
      "A045" ~ round(satisfaction, 1),
      "A042" ~ round(competence, 1),
      "A061" ~ round(continuance, 1),
      
      .default = NA
    )
  ) %>%
  select(CASE, qualitative_question, question_mean, expQual) %>%
  arrange(qualitative_question) %>%
  kable(caption = "Qualitative Experience Question") %>%
  kable_styling(bootstrap_options = "striped")
```



## Purpose of study

```{r}

df %>%
  select(CASE, purpose) %>%
  kable(caption = "Purpose Question") %>%
  kable_styling(bootstrap_options = "striped")
```

## Optional Debrief Question

```{r}
df %>%
  filter(!is.na(debrief_question)) %>%
  select(CASE, debrief_question) %>%
  kable(caption = "Debrief Question") %>%
  kable_styling(bootstrap_options = "striped")
```

