---
title: "Benchmark Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 5)

# Plots Path
output_dir <- "/Users/maxeder/Documents/GitHub/ma_thesis_manuscript/Figures"

# 
# file.path(output_dir, "")
```

# Packages

```{r packages}
library(jsonlite)
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
library(lme4)
library(lmerTest)
library(emmeans)
# library(boot)
```


# Colors

```{r}
colors <- c("Pro" = "#0072B2", "Con" = "#D55E00")

```



# Data Loading & Cleaning

```{r load-data}

unp_data_raw <- fromJSON("../judge_output_final/unprompted_judge_results_multiturn_gemini-2.5-flash.json")
syc_data_raw <- fromJSON("../judge_output_final/sycophantic_judge_results_multiturn_gemini-2.5-flash.json")
cont_data_raw <- fromJSON("../judge_output_final/contrarian_judge_results_multiturn_gemini-2.5-flash.json")


df_unp <- as_tibble(unp_data_raw) %>%
  select(case_id, topic_key, user_stance, topic_key, classified_response) %>%
  unnest(classified_response) %>% 
  select(case_id, topic_key, user_stance, topic_key, turn, mean_score)


df_syc <- as_tibble(syc_data_raw) %>%
  select(case_id, topic_key, user_stance, topic_key, classified_response) %>%
  unnest(classified_response) %>% 
  select(case_id, topic_key, user_stance, topic_key, turn, mean_score)


df_cont <- as_tibble(cont_data_raw) %>%
  select(case_id, topic_key, user_stance, topic_key, classified_response) %>%
  unnest(classified_response) %>%
  select(case_id, topic_key, user_stance, topic_key, turn, mean_score)
```


# Unprompted Plot

```{r}

unp_plot <- ggplot(df_unp, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) +
  
  # Data Points
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.1, size = 1) +
  # geom_boxplot(aes(group = interaction(turn, user_stance)),
  #              position = position_dodge(width = 0.6),
  #              width = 0.4, alpha = 0.2, outlier.size = 0.8, outlier.alpha = 0.3) +
  

  
  # Error Ribbon
  # with standard error
  # stat_summary(geom = "ribbon", fun.data = mean_se, alpha = 0.2, color = NA) +
  # with 95% Confidence Interval
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  
  # Trend Line
  # connect mean score of each turn
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  
  # Mean Points
  # dot for the average at each turn
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
  
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-2, 2), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(breaks = unique(df_unp$turn)) +
  
  labs(
#     title = "Model Agreement with Debate Topic over Turns",
    # title = "Evolution of Model Stance over 8 Turns (Unprompted Condition)",
    # subtitle = "Raw response scores (dots) vs. Average trend (line)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") 

unp_plot
ggsave("plots/unprompted_plot.pdf", plot = unp_plot, width=8, height=5)
ggsave("plots/unprompted_plot.png", plot = unp_plot, width=8, height=5)
ggsave(file.path(output_dir, "benchmark_unprompted_plot.pdf"), plot = unp_plot, width=8, height=5)

```


# Sycophantic Plot

```{r}

syc_plot <- ggplot(df_syc, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) +
  
  # Data Points
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.1, size = 1) +
  
  # Error Ribbon
  # with standard error
  # stat_summary(geom = "ribbon", fun.data = mean_se, alpha = 0.2, color = NA) +
  # with 95% Confidence Interval
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  
  # Trend Line
  # connect mean score of each turn
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  
  # Mean Points
  # dot for the average at each turn
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
  
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-2, 2), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(breaks = unique(df_syc$turn)) +
  
  labs(
#     title = "Model Agreement with Debate Topic over Turns",
    # title = "Evolution of Model Stance over 8 Turns (Sycophantic Condition)",
    # subtitle = "Raw response scores (dots) vs. Average trend (line)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  # theme(legend.position = "bottom", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
  # theme(legend.position = "bottom", panel.border = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
    theme(legend.position = "bottom", panel.border = element_blank(), panel.grid.minor = element_blank())


ggsave("plots/sycophantic_plot.pdf", plot = syc_plot, width=8, height=5)
ggsave("plots/sycophantic_plot.png", plot = syc_plot, width=8, height=5)
ggsave(file.path(output_dir, "benchmark_sycophantic_plot.pdf"), plot = syc_plot, width=8, height=5)
syc_plot

```



# Contrarian Plot

```{r}

cont_plot <- ggplot(df_cont, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) +
  
  # Data Points
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.1, size = 1) +
  
  # Error Ribbon
  # with standard error
  # stat_summary(geom = "ribbon", fun.data = mean_se, alpha = 0.2, color = NA) +
  # with 95% Confidence Interval
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  
  # Trend Line
  # connect mean score of each turn
  stat_summary(fun = mean, geom = "line", linewidth = 1) +

  # Mean Points
  # dot for the average at each turn
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
  
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-2, 2), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(breaks = unique(df_cont$turn)) +
  
  labs(
#     title = "Model Agreement with Debate Topic over Turns",
    # title = "Evolution of Model Stance over 8 Turns (Contrarian Condition)",
    # subtitle = "Raw response scores (dots) vs. Average trend (line)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")


ggsave("plots/contrarian_plot.pdf", plot = cont_plot, width=8, height=5)
ggsave("plots/contrarian_plot.png", plot = cont_plot, width=8, height=5)
ggsave(file.path(output_dir, "benchmark_contrarian_plot.pdf"), plot = cont_plot, width=8, height=5)
cont_plot

```



# Mean of conditions & delta between conditions

```{r}
# remove turn 0
df_unp <- df_unp %>%
  filter(turn!=0)

df_syc <- df_syc %>%
  filter(turn!=0)

df_cont <- df_cont %>%
  filter(turn!=0)

```


```{r}

# unprompted

stance_means_unp <- df_unp %>%
  group_by(user_stance) %>%
  summarize(avg_score = mean(mean_score, na.rm = TRUE))

delta_unp <- stance_means_unp$avg_score[stance_means_unp$user_stance == "Pro"] - 
  stance_means_unp$avg_score[stance_means_unp$user_stance == "Con"]


# sycophancy

stance_means_syc <- df_syc %>%
  group_by(user_stance) %>%
  summarize(avg_score = mean(mean_score, na.rm = TRUE))

delta_syc <- stance_means_syc$avg_score[stance_means_syc$user_stance == "Pro"] - 
  stance_means_syc$avg_score[stance_means_syc$user_stance == "Con"]


# contrarian

stance_means_cont <- df_cont %>%
  group_by(user_stance) %>%
  summarize(avg_score = mean(mean_score, na.rm = TRUE))

delta_cont <- stance_means_cont$avg_score[stance_means_cont$user_stance == "Pro"] - 
  stance_means_cont$avg_score[stance_means_cont$user_stance == "Con"]

# 
# print("Unprompted Condition:")
# print(stance_means_unp)
# print(delta_unp)
# print("Unprompted Condition:")
# print(stance_means_syc)
# print(delta_syc)
# print("Unprompted Condition:")
# print(stance_means_cont)
# print(delta_cont)
```
```{r}

# Combine into one table
comparison <- data.frame(
  Stance = stance_means_unp$user_stance,
  Unprompted = round(stance_means_unp$avg_score, 3),
  Sycophantic = round(stance_means_syc$avg_score, 3),
  Contrarian = round(stance_means_cont$avg_score, 3)
)

# Add a delta row
comparison <- rbind(comparison, data.frame(
  Stance = "Δ (Pro − Con)",
  Unprompted = round(delta_unp, 3),
  Sycophantic = round(delta_syc, 3),
  Contrarian = round(delta_cont, 3)
))

kable(comparison, caption = "Mean Scores by Stance: Unprompted vs. Sycophantic") %>%
  kable_styling(full_width = FALSE) %>%
  row_spec(nrow(comparison), bold = TRUE, background = "#f0f0f0")
```






# Delta Comparison

```{r}


# Combine all conditions
df_all <- bind_rows(
  df_unp %>% mutate(condition = "unprompted"),
  df_syc %>% mutate(condition = "sycophantic"),
  df_cont %>% mutate(condition = "contrarian")
)

df_all$condition <- factor(df_all$condition,
                           levels = c("unprompted", "sycophantic", "contrarian"))
df_all$user_stance <- factor(df_all$user_stance)
df_all$topic <- factor(df_all$topic_key)

# The interaction is your delta
model <- lmer(mean_score ~ condition * user_stance + (1 | topic), data = df_all)

summary(model)
aov_res <- anova(model)

vc <- as.data.frame(VarCorr(model))
topic_var <- vc$vcov[vc$grp == "topic"]
topic_sd <- vc$sdcor[vc$grp == "topic"]

int_numdf <- aov_res["condition:user_stance", "NumDF"]
int_dendf <- aov_res["condition:user_stance", "DenDF"]
int_f <- aov_res["condition:user_stance", "F value"]

# pairwise comparison

emm <- emmeans(model, ~ user_stance | condition)
# The deltas
# pairs(emm)  # pro - con within each condition
deltas <- as.data.frame(pairs(emm))

# Compare deltas across conditions
emm_int <- emmeans(model, ~ condition * user_stance)
interaction_contrasts <- contrast(emm_int, interaction = "pairwise", adjust = "tukey")
# contrast(emm_int, interaction = "pairwise", adjust = "tukey")
interaction_df <- as.data.frame(interaction_contrasts)


interaction_contrasts

# Extract specific values for inline reporting
delta_unp_v <- abs(deltas$estimate[deltas$condition == "unprompted"])
delta_syc_v <- abs(deltas$estimate[deltas$condition == "sycophantic"])
delta_cont_v <- abs(deltas$estimate[deltas$condition == "contrarian"])
```




## Delta Comparison Plot

```{r}
# head(df_all)

# Get emmeans per condition x stance with CIs
emm <- emmeans(model, ~ user_stance | condition)
emm_df <- as.data.frame(emm)

# Compute deltas with CIs from emmeans
delta_emm <- pairs(emm, reverse = TRUE)  # Pro - Con
delta_df <- as.data.frame(delta_emm)
delta_df <- confint(delta_emm) %>% as.data.frame()

# colnames(delta_df)

# order
delta_df$condition <- factor(delta_df$condition, 
                              levels = c("contrarian", "unprompted", "sycophantic"))


# Extract values for inline reporting
delta_syc <- delta_df[delta_df$condition == "sycophantic", ]
delta_unp <- delta_df[delta_df$condition == "unprompted", ]
delta_cnt <- delta_df[delta_df$condition == "contrarian", ]

# Store interaction contrasts
emm_int <- emmeans(model, ~ condition * user_stance)
int_contrasts <- contrast(emm_int, interaction = "pairwise", adjust = "sidak")
int_df <- as.data.frame(int_contrasts)


# delta_df
# Plot
delta_plot <- ggplot(delta_df, aes(x = condition, y = estimate, fill = condition)) +
  geom_col(width = 0.5) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.15) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_x_discrete(labels = c("contrarian" = "Contrarian", 
                               "unprompted" = "Unprompted", 
                               "sycophantic" = "Sycophantic")) +
  labs(
    title = "Stance Shift Toward User (Pro - Con Delta)",
    y = "Delta (Pro - Con)",
    x = NULL
  ) +
  theme_minimal() +
  # theme(legend.position = "none") +
  theme(
    legend.position = "none",
    panel.grid.major.y = element_line(color = "#DDDDDD"),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  # scale_fill_manual(values = c("contrarian" = "#C0392B", "unprompted" = "#2980B9", "sycophantic" = "#27AE60"))
  # scale_fill_manual(values = c("contrarian" = "#7A9AB5", "unprompted" = "#B8B8B8", "sycophantic" = "#9ABB94"))
  scale_fill_manual(values = c("contrarian" = "#D4807A", "unprompted" = "#7BA7C9", "sycophantic" = "#82B88A"))

delta_plot
ggsave("plots/deltas_plot.pdf", plot = delta_plot, width=8, height=5)
ggsave("plots/deltas_plot.png", plot = delta_plot, width=8, height=5)

ggsave(file.path(output_dir, "benchmark_deltas_plot.pdf"), plot = delta_plot, width=8, height=5)

```

<!-- Figure X displays the mean stance shift (Pro – Con delta) with 95% confidence intervals derived from the linear mixed-effects model for each condition. The sycophantic condition exhibited the largest positive shift (δ = `r round(delta_syc$estimate, 2)`, 95% CI [`r round(delta_syc$asymp.LCL, 2)`, `r round(delta_syc$asymp.UCL, 2)`]), indicating strong alignment with the user's stated position. The unprompted condition also showed a positive shift (δ = `r round(delta_unp$estimate, 2)`, 95% CI [`r round(delta_unp$asymp.LCL, 2)`, `r round(delta_unp$asymp.UCL, 2)`]), suggesting a baseline tendency toward agreement even without explicit sycophantic instruction. The contrarian condition produced a negative shift (δ = `r round(delta_cnt$estimate, 2)`, 95% CI [`r round(delta_cnt$asymp.LCL, 2)`, `r round(delta_cnt$asymp.UCL, 2)`]), confirming that the model actively opposed the user's position. All pairwise differences between conditions were significant (*p* < .001, Šidák-corrected). Confidence intervals were estimated analytically from the model's variance-covariance matrix. -->


## Reporting

<!-- A linear mixed-effects model was fit predicting mean stance score from condition (unprompted, sycophantic, contrarian), user stance (pro, con), and their interaction, with topic as a random intercept. The condition × user stance interaction was significant (*F*(`r int_numdf`, `r int_dendf`) = `r format(round(int_f, 1), big.mark = ",")`, *p* < .001) (F(2, 4790) = 12,848.0, p < .001), indicating that the magnitude of stance shift varied across conditions (see Figure X). The sycophantic condition produced the largest shift toward the user's stated position (δ = 2.64, 95% CI [2.60, 2.69]), followed by the unprompted condition (δ = 1.21, 95% CI [1.16, 1.26]), suggesting a baseline tendency toward agreement even without explicit sycophantic instruction. The contrarian condition showed a reversal, with the model shifting against the user's position (δ = −2.50, 95% CI [−2.54, −2.45]). Pairwise interaction contrasts confirmed that all three conditions differed significantly from each other (all p < .001, Šidák-corrected), with the sycophantic–contrarian difference being the largest (Δδ = 5.14), followed by the unprompted–contrarian difference (Δδ = 3.71) and the sycophantic–unprompted difference (Δδ = 1.43). The random intercept for topic accounted for a variance of 0.10 (SD = 0.32), indicating moderate variability across topics. -->


A linear mixed-effects model was fit predicting mean stance score from condition (unprompted, sycophantic, contrarian), user stance (pro, con), and their interaction, with topic as a random intercept. The condition × user stance interaction was significant (*F*(`r int_numdf`, `r int_dendf`) = `r format(round(int_f, 1), big.mark = ",")`, *p* < .001), indicating that the magnitude of stance shift varied across conditions. The sycophantic condition exhibited the largest positive shift (δ = `r round(delta_syc$estimate, 2)`, 95% CI [`r round(delta_syc$asymp.LCL, 2)`, `r round(delta_syc$asymp.UCL, 2)`]), indicating strong alignment with the user's stated position. The unprompted condition also showed a positive shift (δ = `r round(delta_unp$estimate, 2)`, 95% CI [`r round(delta_unp$asymp.LCL, 2)`, `r round(delta_unp$asymp.UCL, 2)`]), suggesting a baseline tendency toward agreement even without explicit sycophantic instruction. The contrarian condition showed a reversal, with the model shifting against the user's position (δ = `r round(delta_cnt$estimate, 2)`, 95% CI [`r round(delta_cnt$asymp.LCL, 2)`, `r round(delta_cnt$asymp.UCL, 2)`]).
Figure X shows the mean stance shift (Pro – Con delta) with 95% confidence intervals for each condition, derived from the linear mixed-effects model.
Pairwise interaction contrasts confirmed that all three conditions differed significantly from each other (all p < .001, Šidák-corrected), with the sycophantic–contrarian difference being the largest (Δδ = `r round(abs(interaction_df$estimate[3]), 2)`), followed by the unprompted–contrarian difference (Δδ = `r round(abs(interaction_df$estimate[2]), 2)`) and the sycophantic–unprompted difference (Δδ = `r round(abs(interaction_df$estimate[1]), 2)`). The random intercept for topic accounted for a variance of `r round(topic_var, 2)` (SD = `r round(topic_sd, 2)`), indicating moderate variability across topics.





<!-- Within each condition, the pro–con difference (δ) was computed as a measure of the model's tendency to shift toward the user's stated position. The sycophantic condition produced the largest shift (δ = `r round(delta_syc_v, 2)`, *z* = `r round(abs(deltas$z.ratio[deltas$condition == "sycophantic"]), 2)`, *p* < .001), followed by the unprompted condition (δ = `r round(delta_unp_v, 2)`, *z* = `r round(abs(deltas$z.ratio[deltas$condition == "unprompted"]), 2)`, *p* < .001). The contrarian condition showed a reversal, with the model shifting against the user's position (δ = `r round(-delta_cont_v, 2)`, *z* = `r round(abs(deltas$z.ratio[deltas$condition == "contrarian"]), 2)`, *p* < .001). -->

<!-- Pairwise interaction contrasts confirmed that all three conditions differed significantly from each other (all *p* < .001, Tukey-adjusted). The sycophantic–contrarian difference was the largest (Δδ = `r round(abs(interaction_df$estimate[3]), 2)`), followed by the unprompted–contrarian difference (Δδ = `r round(abs(interaction_df$estimate[2]), 2)`) and the sycophantic–unprompted difference (Δδ = `r round(abs(interaction_df$estimate[1]), 2)`). -->

<!-- <!-- The random intercept for topic accounted for a variance of 0.10 (SD = 0.32) ??, indicating moderate variability across topics. --> -->

<!-- The random intercept for topic accounted for a variance of `r round(topic_var, 2)` (SD = `r round(topic_sd, 2)`), indicating moderate variability across topics. -->

