---
title: "User Study Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 5)


# Plots Path
output_dir <- "/Users/maxeder/Documents/GitHub/ma_thesis_manuscript/Figures"

# 
# file.path(output_dir, "")
```


# Packages

```{r packages}
library(jsonlite)
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
library(lme4)
library(lmerTest)
library(emmeans)

```
# Colors

```{r}
colors <- c("Pro" = "#0072B2", "Con" = "#D55E00", "Neutral" = "#adadad")

```

# Data Loading & Structuring

```{r load-data}

user_study_raw <- fromJSON("../user_study_data/user_study_judged.json")

df_u_raw <- as_tibble(user_study_raw) %>%
  select(case_id, condition, topic, pre_opinion, classified_response) %>%
  unnest(classified_response) %>% 
  select(case_id, condition, topic, pre_opinion, turn, role, mean_score)


df_u_raw <- df_u_raw %>%
  mutate(
    user_stance = factor(case_when(
      pre_opinion == 4 ~ "Neutral",
      pre_opinion > 4 ~ "Pro",
      pre_opinion < 4 ~ "Con",
    ), levels = c("Con", "Neutral", "Pro"))
  ) 


```

# Turn Histogram

```{r}
pdf(file.path(output_dir, "turn_histogramm.pdf"), width = 8, height = 5)

hist(df_u_raw$turn,
     breaks = "Scott",
     main = "",
     xlab = "Turn",
     ylab = "Frequency",
     xlim = c(0, 20),
     col = "grey80",
     border = "white",
     las = 1,
     cex.main = 1.3,
     cex.lab = 1.1,
     cex.axis = 0.9,
     font.main = 1)
box(bty = "l")

dev.off()

hist(df_u_raw$turn,
     breaks = "Scott",
     main = "",
     xlab = "Turn",
     ylab = "Frequency",
     xlim = c(0, 20),
     col = "grey80",
     border = "white",
     las = 1,
     cex.main = 1.3,
     cex.lab = 1.1,
     cex.axis = 0.9,
     font.main = 1)
```


# Define turn cutoff point
```{r}
quantiles <- quantile(df_u_raw$turn, probs = c(0.5, 0.75, 0.9, 0.95))
print(quantiles)


hist(df_u_raw$turn)

df_u_raw <- df_u_raw %>% filter(turn <= 5)
hist(df_u_raw$turn)
```


```{r split-data}
# By conditions

df_u_unp <- df_u_raw %>% 
  filter(condition == "unprompted") 

df_u_syc <- df_u_raw %>% 
  filter(condition == "sycophantic") 

df_u_cont <- df_u_raw %>% 
  filter(condition == "contrarian") 


# # Assistant turns
# # & remove Neutral
# 
# df_u_unp_assistant <- df_u_unp %>% 
#   filter(role == "assistant") %>% 
#   filter(user_stance != "Neutral") 
# 
# df_u_syc_assistant <- df_u_syc %>% 
#   filter(role == "assistant") %>% 
#   filter(user_stance != "Neutral") 
# 
# df_u_cont_assistant <- df_u_cont %>% 
#   filter(role == "assistant") %>% 
#   filter(user_stance != "Neutral") 
# 
# 
# 
# # User turns
# # & remove Neutral
# 
# 
# df_u_unp_user <- df_u_unp %>% 
#   filter(role == "user") %>%
#   filter(user_stance != "Neutral") 
# 
# df_u_syc_user <- df_u_syc %>% 
#   filter(role == "user") %>%
#   filter(user_stance != "Neutral") 
# 
# df_u_cont_user <- df_u_cont %>% 
#   filter(role == "user") %>%
#   filter(user_stance != "Neutral") 



# Assistant turns

df_u_unp_assistant <- df_u_unp %>% 
  filter(role == "assistant") 

df_u_syc_assistant <- df_u_syc %>% 
  filter(role == "assistant") 

df_u_cont_assistant <- df_u_cont %>% 
  filter(role == "assistant") 



# User turns

df_u_unp_user <- df_u_unp %>% 
  filter(role == "user") 

df_u_syc_user <- df_u_syc %>% 
  filter(role == "user") 

df_u_cont_user <- df_u_cont %>% 
  filter(role == "user") 




  
```



# Assistant Turns Plots

```{r}

unp_plot_assistant <- ggplot(df_u_unp_assistant, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1, shape = 16) +
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
  
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-3, 3), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(limits = c(1, 5)) +
  labs(
    # title = "Evolution of Model Stance over 8 Turns (Unprompted Condition)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") 


syc_plot_assistant <- ggplot(df_u_syc_assistant, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1, shape = 16) +
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
  
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-3, 3), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(limits = c(1, 5)) +
  labs(
    # title = "Evolution of Model Stance over 8 Turns (Sycophantic Condition)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") 

cont_plot_assistant <- ggplot(df_u_cont_assistant, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1, shape = 16) +
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
  
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-3, 3), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(limits = c(1, 5)) +
  labs(
    # title = "Evolution of Model Stance over 8 Turns (Contrarian Condition)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") 

# unp_plot_assistant
# syc_plot_assistant
# cont_plot_assistant
# ggsave("plots/user_study_assistant_unprompted_plot.pdf", plot = unp_plot_assistant, width=8, height=5)
# ggsave("plots/user_study_assistant_sycophantic_plot.pdf", plot = syc_plot_assistant, width=8, height=5)
# ggsave("plots/user_study_assistant_contrarian_plot.pdf", plot = cont_plot_assistant, width=8, height=5)
# 
# 
# ggsave(file.path(output_dir, "user_study_assistant_unprompted_plot.pdf"), plot = unp_plot_assistant, width=8, height=5)
# ggsave(file.path(output_dir, "user_study_assistant_sycophantic_plot.pdf"), plot = syc_plot_assistant, width=8, height=5)
# ggsave(file.path(output_dir, "user_study_assistant_contrarian_plot.pdf"), plot = cont_plot_assistant, width=8, height=5)
# 
# ggsave(file.path(output_dir, "user_study_assistant_unprompted_plot.png"), plot = unp_plot_assistant, width=8, height=5)
# ggsave(file.path(output_dir, "user_study_assistant_sycophantic_plot.png"), plot = syc_plot_assistant, width=8, height=5)
# ggsave(file.path(output_dir, "user_study_assistant_contrarian_plot.png"), plot = cont_plot_assistant, width=8, height=5)


```


## Assistant Delta Comparisons

## Clean Data
```{r}


# Combine all conditions
df_all_assistant <- bind_rows(
  df_u_unp_assistant %>% mutate(condition = "unprompted"),
  df_u_syc_assistant %>% mutate(condition = "sycophantic"),
  df_u_cont_assistant %>% mutate(condition = "contrarian")
)

# remove turn 0
df_all_assistant <- df_all_assistant %>%
  filter(turn!=0)

df_all_assistant

df_all_assistant$condition <- factor(df_all_assistant$condition,
                           levels = c("unprompted", "sycophantic", "contrarian"))
df_all_assistant$user_stance <- factor(df_all_assistant$user_stance)
df_all_assistant$topic <- factor(df_all_assistant$topic)

```


## Models
```{r}


model_assistant <- lmer(mean_score ~ condition * user_stance + (1 | topic), data = df_all_assistant)

summary(model_assistant)
aov_res_assistant <- anova(model_assistant)

aov_res_assistant

vc_a <- as.data.frame(VarCorr(model_assistant))
topic_var_a <- vc_a$vcov[vc_a$grp == "topic"]
topic_sd_a <- vc_a$sdcor[vc_a$grp == "topic"]

int_numdf_a <- aov_res_assistant["condition:user_stance", "NumDF"]
int_dendf_a <- aov_res_assistant["condition:user_stance", "DenDF"]
int_f_a <- aov_res_assistant["condition:user_stance", "F value"]

# pairwise comparison

emm_a <- emmeans(model_assistant, ~ user_stance | condition)
# The deltas
# pairs(emm)  # pro - con within each condition
deltas_a <- as.data.frame(pairs(emm_a))

# Compare deltas across conditions
emm_int_a <- emmeans(model_assistant, ~ condition * user_stance)
interaction_contrasts_a <- contrast(emm_int_a, interaction = "pairwise", adjust = "tukey")
# contrast(emm_int, interaction = "pairwise", adjust = "tukey")
interaction_df_a <- as.data.frame(interaction_contrasts_a)


interaction_contrasts_a

# Extract specific values for inline reporting
delta_unp_a_v <- abs(deltas_a$estimate[deltas_a$condition == "unprompted"])
delta_syc_a_v <- abs(deltas_a$estimate[deltas_a$condition == "sycophantic"])
delta_cont_a_v <- abs(deltas_a$estimate[deltas_a$condition == "contrarian"])



# CIs for inline reporting
delta_emm_a <- pairs(emm_a, reverse = TRUE)
delta_df_a <- as.data.frame(confint(delta_emm_a))

delta_syc_a <- delta_df_a[delta_df_a$condition == "sycophantic", ]
delta_unp_a <- delta_df_a[delta_df_a$condition == "unprompted", ]
delta_cnt_a <- delta_df_a[delta_df_a$condition == "contrarian", ]

# Unprompted significance test
coefs_a <- summary(model_assistant)$coefficients
unp_t_a <- coefs_a["user_stancePro", "t value"]
unp_p_a <- coefs_a["user_stancePro", "Pr(>|t|)"]

```


A linear mixed-effects model was fit to the assistant responses of the real-world conversation data, predicting mean stance score from condition, user stance, and their interaction, with topic as a random intercept. The condition × user stance interaction was significant (*F*(`r int_numdf_a`, `r round(int_dendf_a, 1)`) = `r round(int_f_a, 1)`,*p* < .001), indicating that stance shift varied across conditions. 
However, unlike the benchmark results, the unprompted condition did not show a significant pro–con difference (δ = `r round(delta_unp_a$estimate, 2)`, 95% CI [`r round(delta_unp_a$lower.CL, 2)`, `r round(delta_unp_a$upper.CL, 2)`], *p* = `r round(unp_p_a, 3)`), suggesting that baseline sycophancy was not evident in real-world multi-turn conversations. The sycophantic condition produced a significant positive shift (δ = `r round(delta_syc_a$estimate, 2)`, 95% CI [`r round(delta_syc_a$lower.CL, 2)`, `r round(delta_syc_a$upper.CL, 2)`]), and the contrarian condition produced a significant negative shift (δ = `r round(delta_cnt_a$estimate, 2)`, 95% CI [`r round(delta_cnt_a$lower.CL, 2)`, `r round(delta_cnt_a$upper.CL, 2)`]). Pairwise interaction contrasts confirmed that all three conditions differed significantly from each other (all *p* < .001, Šidák-corrected). The random intercept for topic accounted for a variance of `r round(topic_var_a, 2)` (SD = `r round(topic_sd_a, 2)`).

















<!-- # User Turns Plots -->

<!-- ```{r} -->

<!-- unp_plot_user <- ggplot(df_u_unp_user, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) + -->
<!--   geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1) + -->
<!--   stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) + -->
<!--   stat_summary(fun = mean, geom = "line", linewidth = 1) + -->
<!--   stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) + -->

<!--   scale_color_manual(values = colors) + -->
<!--   scale_fill_manual(values = colors) + -->

<!--   scale_y_continuous(limits = c(-3, 3), breaks = seq(-2, 2, 1)) + -->
<!--   scale_x_continuous(limits = c(1, 5)) + -->
<!--   labs( -->
<!--     # title = "Evolution of User Stance over 8 Turns (Unprompted Condition)", -->
<!--     # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.", -->
<!--     color = "User Stance",  -->
<!--     fill = "User Stance", -->
<!--     x = "Conversation Turn", -->
<!--     y = "Classification Score (Alignment with Topic)" -->
<!--   ) + -->
<!--   theme_minimal() + -->
<!--   theme(legend.position = "bottom")  -->


<!-- syc_plot_user <- ggplot(df_u_syc_user, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) + -->
<!--   geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1) + -->
<!--   stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) + -->
<!--   stat_summary(fun = mean, geom = "line", linewidth = 1) + -->
<!--   stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) + -->

<!--   scale_color_manual(values = colors) + -->
<!--   scale_fill_manual(values = colors) + -->

<!--   scale_y_continuous(limits = c(-3, 3), breaks = seq(-2, 2, 1)) + -->
<!--   scale_x_continuous(limits = c(1, 5)) + -->
<!--   labs( -->
<!--     # title = "Evolution of User Stance over 8 Turns (Sycophantic Condition)", -->
<!--     # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.", -->
<!--     color = "User Stance",  -->
<!--     fill = "User Stance", -->
<!--     x = "Conversation Turn", -->
<!--     y = "Classification Score (Alignment with Topic)" -->
<!--   ) + -->
<!--   theme_minimal() + -->
<!--   theme(legend.position = "bottom")  -->

<!-- cont_plot_user <- ggplot(df_u_cont_user, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) + -->
<!--   geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1) + -->
<!--   stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) + -->
<!--   stat_summary(fun = mean, geom = "line", linewidth = 1) + -->
<!--   stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) + -->

<!--   scale_color_manual(values = colors) + -->
<!--   scale_fill_manual(values = colors) + -->

<!--   scale_y_continuous(limits = c(-3, 3), breaks = seq(-2, 2, 1)) + -->
<!--   scale_x_continuous(limits = c(1, 5)) + -->
<!--   labs( -->
<!--     # title = "Evolution of User Stance over 8 Turns (Contrarian Condition)", -->
<!--     # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.", -->
<!--     color = "User Stance",  -->
<!--     fill = "User Stance", -->
<!--     x = "Conversation Turn", -->
<!--     y = "Classification Score (Alignment with Topic)" -->
<!--   ) + -->
<!--   theme_minimal() + -->
<!--   theme(legend.position = "bottom")  -->

<!-- unp_plot_user -->
<!-- syc_plot_user -->
<!-- cont_plot_user -->

<!-- ggsave("plots/user_study_user_unprompted.pdf", plot = unp_plot_user, width=8, height=5) -->
<!-- ggsave("plots/user_study_user_sycophantic_plot.pdf", plot = syc_plot_user, width=8, height=5) -->
<!-- ggsave("plots/user_study_user_contrarian_plot.pdf", plot = cont_plot_user, width=8, height=5) -->

<!-- ``` -->


# User Turns Plots
```{r}

df_all_user <- bind_rows(
  df_u_unp_user %>% mutate(condition = "unprompted"),
  df_u_syc_user %>% mutate(condition = "sycophantic"),
  df_u_cont_user %>% mutate(condition = "contrarian")
)


df_all_user

ggplot(df_all_user, aes(x = turn, y = mean_score, color = condition, fill = condition)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1, shape = 16) +
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
  
  # scale_color_manual(values = colors) +
  # scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-3, 3), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(limits = c(1, 5)) +
  labs(
    # title = "Evolution of User Stance over 8 Turns (Unprompted Condition)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") 


model_user <- lm(mean_score ~ condition, data = df_all_user)
summary(model_user)

m <- lm(mean_score ~ condition + user_stance, data = df_all_user)
summary(m)
```


```{r}
df_all_user$condition <- relevel(factor(df_all_user$condition), ref = "unprompted")

# df_all_user <- df_all_user %>% filter(turn <= 5)

model <- lmer(mean_score ~ turn * condition + (1 | case_id), data = df_all_user)
summary(model)

# Get predicted values from the model
df_all_user$predicted <- predict(model)

# Condition-level predicted slopes (fixed effects only)
df_all_user$fitted_fixed <- predict(model, re.form = NA)

user_study_user_agreement_plot <- df_all_user  %>%
  mutate(
    condition = factor(condition, levels = c("contrarian", "unprompted", "sycophantic"),
    labels = c("Contrarian", "Unprompted", "Sycophantic"))
  ) %>%
ggplot(aes(x = turn, y = mean_score, colour = condition)) +
  # Individual trajectories (faint)
  # geom_line(aes(group = case_id), alpha = 0.15) +
  # Model-predicted slope per condition (bold)
  geom_jitter(width = 0.15, height = 0.15, alpha = 0.5, size = 1, shape = 16) +
  # stat_summary(aes(y = fitted_fixed), fun = mean, geom = "line", linewidth = 1.5) +
  geom_smooth(method = "lm",  alpha = 0.15) +
  scale_color_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
  # stat_summary(fun.data = "mean_cl_normal", size = 0.4) +
  labs(
    x = "Turn",
    y = "Mean Agreement Score",
    colour = "Condition"
  ) +
  # scale_x_continuous(limits = c(1, 5)) +
  guides(color = "none") +
    theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  ) +
  facet_grid(~ condition) 


user_study_user_agreement_plot
ggsave(file.path(output_dir, "user_study_user_agreement_plot.pdf"), plot = user_study_user_agreement_plot, width=8, height=5)

ggsave(file.path(output_dir, "user_study_user_agreement_plot.png"), plot = user_study_user_agreement_plot, width=8, height=5)
```




```{r}
df_all_user
# model_user <- lm(mean_score ~ condition + pre_opinion, data = df_all_user)
model_user <- lmer(mean_score ~ turn * condition + (1 + turn | participant), data = df_all_user)
summary(model_user)

df_all_user$condition <- relevel(factor(df_all_user$condition), ref = "unprompted")

model <- lmer(mean_score ~ turn * condition + (1 | case_id), data = df_all_user)
summary(model)

df_all_user %>%
ggplot(aes(y = mean_score, x = pre_opinion, color = condition)) +
  geom_jitter(alpha = 0.5, size = 1, color = "black", width = 0.3, shape =16) +
  geom_smooth(method = "lm", alpha = 0.15) +
  stat_summary(fun.data = "mean_cl_normal", size = 0.4) +
  scale_color_manual(values = c("Contrarian" = "#D4807A", "Unprompted" = "#7BA7C9", "Sycophantic" = "#82B88A")) +
  facet_grid(~ condition) +
  guides(color = "none") +
  labs(x='Pre-Interaction Opinion (Centered)', y='Mean Turn Score (Centered)') +
  ylim(-3.3,3.3) +
  theme_minimal(base_size = 13) +
  theme(
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```


<!-- ```{r} -->

<!-- # Combine all conditions -->
<!-- df_all_user <- bind_rows( -->
<!--   df_u_unp_user %>% mutate(condition = "unprompted"), -->
<!--   df_u_syc_user %>% mutate(condition = "sycophantic"), -->
<!--   df_u_cont_user %>% mutate(condition = "contrarian") -->
<!-- ) -->


<!-- df_all_user -->

<!-- df_all_user$condition <- factor(df_all_user$condition, -->
<!--                            levels = c("unprompted", "sycophantic", "contrarian")) -->
<!-- df_all_user$user_stance <- factor(df_all_user$user_stance) -->
<!-- df_all_user$topic <- factor(df_all_user$topic) -->

<!-- ``` -->


<!-- ## Models -->
<!-- ```{r} -->


<!-- model_user <- lmer(mean_score ~ condition * user_stance + (1 | topic), data = df_all_user) -->

<!-- summary(model_user) -->
<!-- aov_res_user <- anova(model_user) -->

<!-- aov_res_user -->

<!-- vc_u <- as.data.frame(VarCorr(model_user)) -->
<!-- topic_var_u <- vc_a$vcov[vc_a$grp == "topic"] -->
<!-- topic_sd_u <- vc_a$sdcor[vc_a$grp == "topic"] -->

<!-- int_numdf_u <- aov_res_user["condition:user_stance", "NumDF"] -->
<!-- int_dendf_u <- aov_res_user["condition:user_stance", "DenDF"] -->
<!-- int_f_u <- aov_res_user["condition:user_stance", "F value"] -->

<!-- # pairwise comparison -->

<!-- emm_u <- emmeans(model_user, ~ user_stance | condition) -->
<!-- # pro - con within each condition -->
<!-- deltas_u <- as.data.frame(pairs(emm_u)) -->

<!-- # Compare deltas across conditions -->
<!-- emm_int_u <- emmeans(model_user, ~ condition * user_stance) -->
<!-- interaction_contrasts_u <- contrast(emm_int_u, interaction = "pairwise", adjust = "tukey") -->
<!-- interaction_df_u <- as.data.frame(interaction_contrasts_u) -->


<!-- interaction_contrasts_u -->

<!-- # Extract specific values for inline reporting -->
<!-- delta_unp_u_v <- abs(deltas_u$estimate[deltas_u$condition == "unprompted"]) -->
<!-- delta_syc_u_v <- abs(deltas_u$estimate[deltas_u$condition == "sycophantic"]) -->
<!-- delta_cont_u_v <- abs(deltas_u$estimate[deltas_u$condition == "contrarian"]) -->



<!-- # CIs for inline reporting -->
<!-- delta_emm_u <- pairs(emm_a, reverse = TRUE) -->
<!-- delta_df_u <- as.data.frame(confint(delta_emm_u)) -->

<!-- delta_syc_u <- delta_df_u[delta_df_u$condition == "sycophantic", ] -->
<!-- delta_unp_u <- delta_df_u[delta_df_u$condition == "unprompted", ] -->
<!-- delta_cnt_u <- delta_df_u[delta_df_u$condition == "contrarian", ] -->

<!-- # Unprompted significance test -->
<!-- coefs_u <- summary(model_assistant)$coefficients -->
<!-- unp_t_u <- coefs_u["user_stancePro", "t value"] -->
<!-- unp_p_u <- coefs_u["user_stancePro", "Pr(>|t|)"] -->

<!-- ``` -->



