---
title: "User Study Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 5)
```


# Packages

```{r packages}
library(jsonlite)
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
library(lme4)
library(lmerTest)
library(emmeans)

```
# Colors

```{r}
colors <- c("Pro" = "#0072B2", "Con" = "#D55E00")

```

# Data Loading & Structuring

```{r load-data}

user_study_raw <- fromJSON("../user_study_data/user_study_judged.json")

df_u_raw <- as_tibble(user_study_raw) %>%
  select(case_id, condition, topic, pre_opinion, classified_response) %>%
  unnest(classified_response) %>% 
  select(case_id, condition, topic, pre_opinion, turn, role, mean_score)





df_u_raw <- df_u_raw %>%
  mutate(
    user_stance = factor(case_when(
      pre_opinion == 4 ~ "Neutral",
      pre_opinion > 4 ~ "Pro",
      pre_opinion < 4 ~ "Con",
    ), levels = c("Con", "Neutral", "Pro"))
  ) 



# By conditions

df_u_unp <- df_u_raw %>% 
  filter(condition == "unprompted") 

df_u_syc <- df_u_raw %>% 
  filter(condition == "sycophantic") 

df_u_cont <- df_u_raw %>% 
  filter(condition == "contrarian") 


# Assistant turns
# & remove Neutral

df_u_unp_assistant <- df_u_unp %>% 
  filter(role == "assistant") %>% 
  filter(user_stance != "Neutral") 

df_u_syc_assistant <- df_u_syc %>% 
  filter(role == "assistant") %>% 
  filter(user_stance != "Neutral") 

df_u_cont_assistant <- df_u_cont %>% 
  filter(role == "assistant") %>% 
  filter(user_stance != "Neutral") 



# User turns
# & remove Neutral


df_u_unp_user <- df_u_unp %>% 
  filter(role == "user") %>%
  filter(user_stance != "Neutral") 

df_u_syc_user <- df_u_syc %>% 
  filter(role == "user") %>%
  filter(user_stance != "Neutral") 

df_u_cont_user <- df_u_cont %>% 
  filter(role == "user") %>%
  filter(user_stance != "Neutral") 


```


# Assistant Turns Plots

```{r}

unp_plot_assistant <- ggplot(df_u_unp_assistant, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1) +
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
  
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-2, 2), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(limits = c(1, 8)) +
  labs(
    # title = "Evolution of Model Stance over 8 Turns (Unprompted Condition)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") 


syc_plot_assistant <- ggplot(df_u_syc_assistant, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1) +
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
  
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-2, 2), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(limits = c(1, 8)) +
  labs(
    # title = "Evolution of Model Stance over 8 Turns (Sycophantic Condition)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") 

cont_plot_assistant <- ggplot(df_u_cont_assistant, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1) +
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
  
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-2, 2), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(limits = c(1, 8)) +
  labs(
    # title = "Evolution of Model Stance over 8 Turns (Contrarian Condition)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") 

unp_plot_assistant
syc_plot_assistant
cont_plot_assistant
ggsave("plots/user_study_assistant_unprompted_plot.pdf", plot = unp_plot_assistant, width=8, height=5)
ggsave("plots/user_study_assistant_sycophantic_plot.pdf", plot = syc_plot_assistant, width=8, height=5)
ggsave("plots/user_study_assistant_contrarian_plot.pdf", plot = cont_plot_assistant, width=8, height=5)

```


## Assistant Delta Comparisons

## Clean Data
```{r}


# Combine all conditions
df_all_assistant <- bind_rows(
  df_u_unp_assistant %>% mutate(condition = "unprompted"),
  df_u_syc_assistant %>% mutate(condition = "sycophantic"),
  df_u_cont_assistant %>% mutate(condition = "contrarian")
)

# remove turn 0
df_all_assistant <- df_all_assistant %>%
  filter(turn!=0)

df_all_assistant

df_all_assistant$condition <- factor(df_all_assistant$condition,
                           levels = c("unprompted", "sycophantic", "contrarian"))
df_all_assistant$user_stance <- factor(df_all_assistant$user_stance)
df_all_assistant$topic <- factor(df_all_assistant$topic)

```


## Models
```{r}



# The interaction is your delta
model_assistant <- lmer(mean_score ~ condition * user_stance + (1 | topic), data = df_all_assistant)

summary(model_assistant)
aov_res_assistant <- anova(model_assistant)

aov_res_assistant

vc_a <- as.data.frame(VarCorr(model_assistant))
topic_var_a <- vc_a$vcov[vc_a$grp == "topic"]
topic_sd_a <- vc_a$sdcor[vc_a$grp == "topic"]

int_numdf_a <- aov_res_assistant["condition:user_stance", "NumDF"]
int_dendf_a <- aov_res_assistant["condition:user_stance", "DenDF"]
int_f_a <- aov_res_assistant["condition:user_stance", "F value"]

# pairwise comparison

emm_a <- emmeans(model_assistant, ~ user_stance | condition)
# The deltas
# pairs(emm)  # pro - con within each condition
deltas_a <- as.data.frame(pairs(emm_a))

# Compare deltas across conditions
emm_int_a <- emmeans(model_assistant, ~ condition * user_stance)
interaction_contrasts_a <- contrast(emm_int_a, interaction = "pairwise", adjust = "tukey")
# contrast(emm_int, interaction = "pairwise", adjust = "tukey")
interaction_df_a <- as.data.frame(interaction_contrasts_a)


interaction_contrasts_a

# Extract specific values for inline reporting
delta_unp_a_v <- abs(deltas_a$estimate[deltas_a$condition == "unprompted"])
delta_syc_a_v <- abs(deltas_a$estimate[deltas_a$condition == "sycophantic"])
delta_cont_a_v <- abs(deltas_a$estimate[deltas_a$condition == "contrarian"])



# CIs for inline reporting
delta_emm_a <- pairs(emm_a, reverse = TRUE)
delta_df_a <- as.data.frame(confint(delta_emm_a))

delta_syc_a <- delta_df_a[delta_df_a$condition == "sycophantic", ]
delta_unp_a <- delta_df_a[delta_df_a$condition == "unprompted", ]
delta_cnt_a <- delta_df_a[delta_df_a$condition == "contrarian", ]

# Unprompted significance test
coefs_a <- summary(model_assistant)$coefficients
unp_t_a <- coefs_a["user_stancePro", "t value"]
unp_p_a <- coefs_a["user_stancePro", "Pr(>|t|)"]

```


A linear mixed-effects model was fit to the assistant responses of the real-world conversation data, predicting mean stance score from condition, user stance, and their interaction, with topic as a random intercept. The condition × user stance interaction was significant (*F*(`r int_numdf_a`, `r round(int_dendf_a, 1)`) = `r round(int_f_a, 1)`,*p* < .001), indicating that stance shift varied across conditions. 
However, unlike the benchmark results, the unprompted condition did not show a significant pro–con difference (δ = `r round(delta_unp_a$estimate, 2)`, 95% CI [`r round(delta_unp_a$lower.CL, 2)`, `r round(delta_unp_a$upper.CL, 2)`], *p* = `r round(unp_p_a, 3)`), suggesting that baseline sycophancy was not evident in real-world multi-turn conversations. The sycophantic condition produced a significant positive shift (δ = `r round(delta_syc_a$estimate, 2)`, 95% CI [`r round(delta_syc_a$lower.CL, 2)`, `r round(delta_syc_a$upper.CL, 2)`]), and the contrarian condition produced a significant negative shift (δ = `r round(delta_cnt_a$estimate, 2)`, 95% CI [`r round(delta_cnt_a$lower.CL, 2)`, `r round(delta_cnt_a$upper.CL, 2)`]). Pairwise interaction contrasts confirmed that all three conditions differed significantly from each other (all *p* < .001, Šidák-corrected). The random intercept for topic accounted for a variance of `r round(topic_var_a, 2)` (SD = `r round(topic_sd_a, 2)`).

















# User Turns Plots

```{r}

unp_plot_user <- ggplot(df_u_unp_user, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1) +
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
  
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-2, 2), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(limits = c(1, 8)) +
  labs(
    # title = "Evolution of User Stance over 8 Turns (Unprompted Condition)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") 


syc_plot_user <- ggplot(df_u_syc_user, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1) +
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
  
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-2, 2), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(limits = c(1, 8)) +
  labs(
    # title = "Evolution of User Stance over 8 Turns (Sycophantic Condition)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") 

cont_plot_user <- ggplot(df_u_cont_user, aes(x = turn, y = mean_score, color = user_stance, fill = user_stance)) +
  geom_jitter(width = 0.05, height = 0.05, alpha = 0.3, size = 1) +
  stat_summary(geom = "ribbon", fun.data = mean_cl_normal, alpha = 0.2, color = NA) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun = mean, geom = "point", size = 2.5, shape = 21, color = "white", stroke = 1) +
    
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  
  scale_y_continuous(limits = c(-2, 2), breaks = seq(-2, 2, 1)) +
  scale_x_continuous(limits = c(1, 8)) +
  labs(
    # title = "Evolution of User Stance over 8 Turns (Contrarian Condition)",
    # subtitle = "Model responses to synthetic pro/con user inputs across 5 debate topics (T=0; 2 replications).\nPoints show mean stance scores calculated from 3 sentences per response. Ribbon shows 95% Confidence Interval.",
    color = "User Stance", 
    fill = "User Stance",
    x = "Conversation Turn",
    y = "Classification Score (Alignment with Topic)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") 

unp_plot_user
syc_plot_user
cont_plot_user

ggsave("plots/user_study_user_unprompted.pdf", plot = unp_plot_user, width=8, height=5)
ggsave("plots/user_study_user_sycophantic_plot.pdf", plot = syc_plot_user, width=8, height=5)
ggsave("plots/user_study_user_contrarian_plot.pdf", plot = cont_plot_user, width=8, height=5)

```








