---
title: "SycLLM Pilot User Study — Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# 1. Setup

```{r packages}
library(tidyverse)
library(psych)
library(ggplot2)
library(knitr)
library(kableExtra)
library(lme4)
library(emmeans)
```

# 2. Data Loading & Cleaning

The raw file is UTF-16 LE tab-separated. We read it with `readr::read_tsv` using the
appropriate encoding, then apply the following exclusion criteria:

- Remove incomplete cases (`FINISHED != 1`)
- Remove admin / test entries (`MODE == "admin"` or `age == "test"` or `age == "0"`)

```{r load-data}

FILE_PATH = "data/pilot_data.csv"
raw <- read_tsv(FILE_PATH, locale = locale(encoding = "UTF-16LE"),  show_col_types = FALSE)

# Inspect
glimpse(raw)
```

```{r clean-data}
df <- raw %>%
  # keep only finished participants
  filter(FINISHED == 1) %>%
  # drop admin / test entries
  filter(MODE != "admin", age != "test", as.numeric(age) > 0) %>%
  mutate(
    age             = as.numeric(age),
    gender          = factor(gender, levels = 1:2, labels = c("male", "female")),
    model_condition = factor(model_condition,
                             levels = 1:3,
                             labels = c("unprompted", "sycophantic", "contrarian")),
    topic  = factor(topic),
    stance = factor(stance)
  )

cat("N after exclusions:", nrow(df), "\n")
```

```{r attitude-scoring}
# Helper: row-wise mean ignoring NA
row_mean <- function(df, cols) {
  rowMeans(df[, cols], na.rm = TRUE)
}

# For each participant the relevant pre/post attitude items depend on their topic.
# We build a long "opinion_pre" / "opinion_post" from the topic-specific columns.
topics <- c("space", "quota", "taxes", "animals", "speech")

df <- df %>%
  mutate(
    # Pre-attitude: mean of the four items for the assigned topic
    opinion_pre = case_when(
      topic == "space"   ~ row_mean(., paste0("space_",   sprintf("%02d", 1:4))),
      topic == "quota"   ~ row_mean(., paste0("quota_",   sprintf("%02d", 1:4))),
      topic == "taxes"   ~ row_mean(., paste0("taxes_",   sprintf("%02d", 1:4))),
      topic == "animals" ~ row_mean(., paste0("animals_", sprintf("%02d", 1:4))),
      topic == "speech"  ~ row_mean(., paste0("speech_",  sprintf("%02d", 1:4)))
    ),
    # Post-attitude: mean of the four post items
    opinion_post = case_when(
      topic == "space"   ~ row_mean(., paste0("post_space_",   sprintf("%02d", 1:4))),
      topic == "quota"   ~ row_mean(., paste0("post_quota_",   sprintf("%02d", 1:4))),
      topic == "taxes"   ~ row_mean(., paste0("post_taxes_",   sprintf("%02d", 1:4))),
      topic == "animals" ~ row_mean(., paste0("post_animals_", sprintf("%02d", 1:4))),
      topic == "speech"  ~ row_mean(., paste0("post_speech_",  sprintf("%02d", 1:4)))
    ),
    opinion_change = opinion_post - opinion_pre,
    # Scale composites
    competence  = row_mean(., c("competence_01", "competence_02", "competence_03")),
    trust       = row_mean(., c("trust_01",       "trust_02",       "trust_03")),
    satisfaction = row_mean(., paste0("satisfaction_0", 1:5)),
    sycophancy  = row_mean(., c("sycophancy_01",  "sycophancy_02")),
    credibility = row_mean(., c("credibility_01", "credibility_02", "credibility_03")),
    continuance = row_mean(., c("continuance_01", "continuance_02"))
  )
```

# 3. Sample Description

## Age

```{r sample-desc}
df %>%
  summarise(
    N        = n(),
    Age_M    = round(mean(age, na.rm = TRUE), 1),
    Age_SD   = round(sd(age, na.rm = TRUE), 1),
    Age_min  = min(age, na.rm = TRUE),
    Age_max  = max(age, na.rm = TRUE)
  ) %>%
  kable(caption = "Sample demographics") %>%
  kable_styling(bootstrap_options = "striped")
```

## Gender

```{r gender-table}
df %>%
  count(gender) %>%
  mutate(pct = round(100 * n / sum(n), 1)) %>%
  kable(caption = "Gender distribution") %>%
  kable_styling(bootstrap_options = "striped")
```

## Model and Topic Conditions

```{r condition-table}
df %>%
  count(model_condition, topic) %>%
  pivot_wider(names_from = topic, values_from = n, values_fill = 0) %>%
  kable(caption = "Participants per condition × topic") %>%
  kable_styling(bootstrap_options = "striped")
```

## Time Overview

```{r time-table}
df %>%
  summarise(
    N        = n(),
    Time_M    = round(round(mean(TIME_SUM, na.rm = TRUE), 1)/60, 1),
    Time_SD   = round(round(sd(TIME_SUM, na.rm = TRUE), 1)/60, 1),
    Time_min  = min(TIME_SUM, na.rm = TRUE)/60,
    Time_max  = max(TIME_SUM, na.rm = TRUE)/60
  ) %>%
  kable(caption = "Sample demographics") %>%
  kable_styling(bootstrap_options = "striped")
```

# 4. Scale Reliability

```{r reliability}
alpha_table <- tibble(
  Scale = c("Competence", "Trust", "Satisfaction", "Sycophancy",
            "Credibility", "Continuance"),
  Items = c(3, 3, 5, 2, 3, 2),
  Alpha = c(
    alpha(df[, c("competence_01", "competence_02", "competence_03")])$total$raw_alpha,
    alpha(df[, c("trust_01", "trust_02", "trust_03")])$total$raw_alpha,
    alpha(df[, paste0("satisfaction_0", 1:5)])$total$raw_alpha,
    alpha(df[, c("sycophancy_01", "sycophancy_02")])$total$raw_alpha,
    alpha(df[, c("credibility_01", "credibility_02", "credibility_03")])$total$raw_alpha,
    alpha(df[, c("continuance_01", "continuance_02")])$total$raw_alpha
  )
) %>%
  mutate(Alpha = round(Alpha, 3))

alpha_table %>%
  kable(caption = "Cronbach's alpha for composite scales") %>%
  kable_styling(bootstrap_options = "striped")
```

# 5. Descriptive Statistics by Condition

```{r descriptives}
df %>%
  group_by(model_condition) %>%
  summarise(
    n            = n(),
    opinion_pre_M    = round(mean(opinion_pre,    na.rm = TRUE), 2),
    opinion_pre_SD   = round(sd(opinion_pre,      na.rm = TRUE), 2),
    opinion_post_M   = round(mean(opinion_post,   na.rm = TRUE), 2),
    opinion_post_SD  = round(sd(opinion_post,     na.rm = TRUE), 2),
    opinion_change_M = round(mean(opinion_change, na.rm = TRUE), 2),
    opinion_change_SD= round(sd(opinion_change,   na.rm = TRUE), 2),
    competence_M = round(mean(competence, na.rm = TRUE), 2),
    trust_M      = round(mean(trust,      na.rm = TRUE), 2),
    satisfaction_M = round(mean(satisfaction, na.rm = TRUE), 2),
    sycophancy_M = round(mean(sycophancy, na.rm = TRUE), 2),
    credibility_M = round(mean(credibility, na.rm = TRUE), 2),
    continuance_M = round(mean(continuance, na.rm = TRUE), 2)
  ) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  kable(caption = "Descriptive statistics by condition") %>%
  kable_styling(bootstrap_options = "striped")
```

# 6. Attitude Change

## 6.1 Pre–Post within Each Condition

```{r att-change-tests}
for (cond in levels(df$model_condition)) {
  sub <- df %>% filter(model_condition == cond, !is.na(opinion_pre), !is.na(opinion_post))
  if (nrow(sub) < 2) next
  t_res <- t.test(sub$opinion_post, sub$opinion_pre, paired = TRUE)
  cat(sprintf(
    "\n%s (n=%d): M_pre=%.2f, M_post=%.2f, diff=%.2f, t(%d)=%.2f, p=%.3f\n",
    cond, nrow(sub),
    mean(sub$opinion_pre), mean(sub$opinion_post), mean(sub$opinion_change),
    t_res$parameter, t_res$statistic, t_res$p.value
  ))
}
```

## 6.2 Attitude Change by Condition

```{r att-change-plot, fig.width=7, fig.height=4}
df_att <- df %>%
  filter(!is.na(opinion_pre), !is.na(opinion_post)) %>%
  pivot_longer(cols = c(opinion_pre, opinion_post),
               names_to = "time", values_to = "attitude") %>%
  mutate(time = factor(time, levels = c("opinion_pre", "opinion_post"),
                       labels = c("Pre", "Post")))

df_att %>%
  ggplot(aes(x = time, y = attitude, colour = model_condition, group = CASE)) +
  geom_line(alpha = 0.4) +
  geom_point(size = 2) +
  stat_summary(aes(group = model_condition), fun = mean,
               geom = "line", linewidth = 1.2, linetype = "dashed") +
  facet_wrap(~model_condition) +
  scale_colour_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "Individual attitude trajectories by condition",
       x = "Time", y = "Attitude (1–5)", colour = "Condition") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

```{r att-change-bar, fig.width=6, fig.height=4}
df %>%
  filter(!is.na(opinion_change)) %>%
  ggplot(aes(x = model_condition, y = opinion_change, fill = model_condition)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.1, size = 2, colour = "black", alpha = 0.6) +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "Attitude change (post – pre) by condition",
       x = "Condition", y = "Attitude change") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

## 6.3 ANOVA / Kruskal-Wallis on Attitude Change

```{r att-change-anova}
df_opinion_change <- df %>% filter(!is.na(opinion_change))

if (nrow(df_opinion_change) >= 5) {
  aov_res <- aov(opinion_change ~ model_condition, data = df_opinion_change)
  print(summary(aov_res))

  # Non-parametric alternative given small N
  kw <- kruskal.test(opinion_change ~ model_condition, data = df_opinion_change)
  cat(sprintf("\nKruskal-Wallis: chi²(%.0f) = %.2f, p = %.3f\n",
              kw$parameter, kw$statistic, kw$p.value))
}
```

# 7. Outcome Measures by Condition

```{r outcomes-long}
outcomes <- c("competence", "trust", "satisfaction", "sycophancy",
              "credibility", "continuance")

df_out <- df %>%
  select(CASE, model_condition, all_of(outcomes)) %>%
  pivot_longer(cols = all_of(outcomes),
               names_to = "outcome", values_to = "value") %>%
  filter(!is.na(value))
```

```{r outcomes-plot, fig.width=9, fig.height=6}
df_out %>%
  ggplot(aes(x = model_condition, y = value, fill = model_condition)) +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.12, size = 1.5, alpha = 0.6) +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  facet_wrap(~outcome, scales = "free_y") +
  labs(title = "Outcome measures by condition",
       x = "Condition", y = "Score", fill = "Condition") +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 25, hjust = 1),
        legend.position = "none")
```

```{r outcomes-kw}
for (out in outcomes) {
  sub <- df %>% filter(!is.na(.data[[out]]))
  if (length(unique(sub$model_condition)) < 2 || nrow(sub) < 4) next
  kw <- kruskal.test(as.formula(paste(out, "~ model_condition")), data = sub)
  cat(sprintf("%-15s  chi²(%.0f)=%.2f  p=%.3f\n",
              out, kw$parameter, kw$statistic, kw$p.value))
}
```

# 8. Perceived Sycophancy

Does perceived sycophancy (`sycophancy_01/02`) differ between conditions?

```{r sycophancy-check}
df %>%
  group_by(model_condition) %>%
  summarise(
    n = n(),
    M = round(mean(sycophancy, na.rm = TRUE), 2),
    SD = round(sd(sycophancy, na.rm = TRUE), 2)
  ) %>%
  kable(caption = "Perceived sycophancy by condition") %>%
  kable_styling(bootstrap_options = "striped")
```

```{r sycophancy-plot, fig.width=5, fig.height=4}
df %>%
  filter(!is.na(sycophancy)) %>%
  ggplot(aes(x = model_condition, y = sycophancy, fill = model_condition)) +
  geom_boxplot(alpha = 0.7, width = 0.4) +
  geom_jitter(width = 0.12, size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("steelblue", "darkorange", "firebrick")) +
  labs(title = "Perceived sycophancy by condition",
       x = "Condition", y = "Sycophancy (1–5)") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

# 9. Correlations Among Outcome Measures

```{r correlations, fig.width=7, fig.height=6}
df_cor <- df %>%
  select(opinion_change, competence, trust, satisfaction,
         sycophancy, credibility, continuance) %>%
  drop_na()

cor_mat <- cor(df_cor, use = "pairwise.complete.obs", method = "spearman")
round(cor_mat, 2) %>%
  kable(caption = "Spearman correlations among outcome measures") %>%
  kable_styling(bootstrap_options = "striped")
```

# 10. Qualitative Results

## Randomized Qualitative Experience Question

```{r qual-results}

labels <- c(
  "A043" = "Trust",
  "A059" = "Sycophancy",
  "A060" = "Credibility",
  "A045" = "Satisfaction",
  "A042" = "Competence",
  "A061" = "Continuance"
)

# df %>%
#   summarise(qualQ = labels[qualQ],
#             if (qualQ == "Competence") competence) %>%
#   kable(caption = "Open Answers") %>%
#   kable_styling(bootstrap_options = "striped")


df %>%
  mutate(
    qualitative_question = labels[qualQ],
    question_mean = case_match(
      qualQ,
      "A043" ~ round(trust, 1),
      "A059" ~ round(sycophancy, 1),
      "A060" ~ round(credibility, 1),
      "A045" ~ round(satisfaction, 1),
      "A042" ~ round(competence, 1),
      "A061" ~ round(continuance, 1),
      
      .default = NA
    )
  ) %>%
  select(CASE, qualitative_question, question_mean, expQual) %>%
  arrange(qualitative_question) %>%
  kable(caption = "Qualitative Experience Question") %>%
  kable_styling(bootstrap_options = "striped")
```



## Purpose of study

```{r}

df %>%
  select(CASE, purpose) %>%
  kable(caption = "Purpose Question") %>%
  kable_styling(bootstrap_options = "striped")
```

## Optional Debrief Question

```{r}
df %>%
  filter(!is.na(debrief_question)) %>%
  select(CASE, debrief_question) %>%
  kable(caption = "Debrief Question") %>%
  kable_styling(bootstrap_options = "striped")
```

